---
title: "chapter 4"
author: "Min-Yao"
date: "2019年4月21日"
output: 
  html_document: 
    keep_md: yes
---
# Geocentric Models

## 4.1. Why normal distributions are normal
### 4.1.1. Normal by addition.

```{r}
library(rethinking)
```


```{r}
## R code 4.1
pos <- replicate( 1000 , sum( runif(16,-1,1) ) )
hist(pos)
plot(density(pos))
dens(pos, norm.comp=TRUE)
```

### 4.1.2. Normal by multiplication.

```{r}
## R code 4.2
prod( 1 + runif(12,0,0.1) )

## R code 4.3
growth <- replicate( 10000 , prod( 1 + runif(12,0,0.1) ) )
dens( growth , norm.comp=TRUE )

## R code 4.4
big <- replicate( 10000 , prod( 1 + runif(12,0,0.5) ) )
dens( big , norm.comp=TRUE )
small <- replicate( 10000 , prod( 1 + runif(12,0,0.01) ) )
dens( small , norm.comp=TRUE )
```

### 4.1.3. Normal by log-multiplication.

```{r}
## R code 4.5
log.big <- replicate( 10000 , log(prod(1 + runif(12,0,0.5))) )
dens( log.big , norm.comp=TRUE )
```

### 4.1.4. Using Gaussian distributions.
## 4.2. A language for describing models
### 4.2.1. Re-describing the globe tossing model.

```{r}
## R code 4.6
w <- 6; n <- 9;
p_grid <- seq(from=0,to=1,length.out=100)
posterior <- dbinom(w,n,p_grid)*dunif(p_grid,0,1)
posterior <- posterior/sum(posterior)
```

## 4.3. A Gaussian model of height

### 4.3.1. The data.

```{r}
## R code 4.7
library(rethinking)
data(Howell1)
d <- Howell1

## R code 4.8
str( d )

## R code 4.9
precis( d )

## R code 4.10
d$height

## R code 4.11
d2 <- d[ d$age >= 18 , ]
precis( d2 )

```

### 4.3.2. The model.

```{r}
dens(d2$height)

## R code 4.12
curve( dnorm( x , 178 , 20 ) , from=100 , to=250 )

## R code 4.13
curve( dunif( x , 0 , 50 ) , from=-10 , to=60 )

## R code 4.14
sample_mu <- rnorm( 1e4 , 178 , 20 )
sample_sigma <- runif( 1e4 , 0 , 50 )
prior_h <- rnorm( 1e4 , sample_mu , sample_sigma )
dens( prior_h )

## R code 4.15
sample_mu <- rnorm( 1e4 , 178 , 100 )
prior_h <- rnorm( 1e4 , sample_mu , sample_sigma )
dens( prior_h )
```

### 4.3.3. Grid approximation of the posterior distribution.

```{r}
## R code 4.16
mu.list <- seq( from=140, to=160 , length.out=200 )
sigma.list <- seq( from=4 , to=9 , length.out=200 )
post <- expand.grid( mu=mu.list , sigma=sigma.list )
post$LL <- sapply( 1:nrow(post) , function(i) sum( dnorm(
                d2$height ,
                mean=post$mu[i] ,
                sd=post$sigma[i] ,
                log=TRUE ) ) )
post$prod <- post$LL + dnorm( post$mu , 178 , 20 , TRUE ) +
    dunif( post$sigma , 0 , 50 , TRUE )
post$prob <- exp( post$prod - max(post$prod) )

## R code 4.17
contour_xyz( post$mu , post$sigma , post$prob )

## R code 4.18
image_xyz( post$mu , post$sigma , post$prob )
```

### 4.3.4. Sampling from the posterior.

```{r}
## R code 4.19
sample.rows <- sample( 1:nrow(post) , size=1e4 , replace=TRUE ,
    prob=post$prob )
sample.mu <- post$mu[ sample.rows ]
sample.sigma <- post$sigma[ sample.rows ]

## R code 4.20
plot( sample.mu , sample.sigma , cex=0.5 , pch=16 , col=col.alpha(rangi2,0.1) )

## R code 4.21
dens( sample.mu )
dens( sample.sigma )

## R code 4.22
HPDI( sample.mu )
HPDI( sample.sigma )

## R code 4.23
d3 <- sample( d2$height , size=20 )

## R code 4.24
mu.list <- seq( from=150, to=170 , length.out=200 )
sigma.list <- seq( from=4 , to=20 , length.out=200 )
post2 <- expand.grid( mu=mu.list , sigma=sigma.list )
post2$LL <- sapply( 1:nrow(post2) , function(i)
    sum( dnorm( d3 , mean=post2$mu[i] , sd=post2$sigma[i] ,
    log=TRUE ) ) )
post2$prod <- post2$LL + dnorm( post2$mu , 178 , 20 , TRUE ) +
    dunif( post2$sigma , 0 , 50 , TRUE )
post2$prob <- exp( post2$prod - max(post2$prod) )
sample2.rows <- sample( 1:nrow(post2) , size=1e4 , replace=TRUE ,
    prob=post2$prob )
sample2.mu <- post2$mu[ sample2.rows ]
sample2.sigma <- post2$sigma[ sample2.rows ]
plot( sample2.mu , sample2.sigma , cex=0.5 ,
    col=col.alpha(rangi2,0.1) ,
    xlab="mu" , ylab="sigma" , pch=16 )

## R code 4.25
dens( sample2.sigma , norm.comp=TRUE )
```

### 4.3.5. Finding the posterior distribution with quap.

```{r}
## R code 4.26
library(rethinking)
data(Howell1)
d <- Howell1
d2 <- d[ d$age >= 18 , ]

## R code 4.27
flist <- alist(
    height ~ dnorm( mu , sigma ) ,
    mu ~ dnorm( 178 , 20 ) ,
    sigma ~ dunif( 0 , 50 )
)

## R code 4.28
m4.1 <- quap( flist , data=d2 )

## R code 4.29
precis( m4.1 )
precis( m4.1, prob = 0.95 )

## R code 4.30
start <- list(
    mu=mean(d2$height),
    sigma=sd(d2$height)
)
m4.1 <- quap( flist , data=d2 , start=start )

## R code 4.31
m4.2 <- quap(
    alist(
        height ~ dnorm( mu , sigma ) ,
        mu ~ dnorm( 178 , 0.1 ) ,
        sigma ~ dunif( 0 , 50 )
    ) , data=d2 )
precis( m4.2 )
```

### 4.3.6. Sampling from a quap. 

```{r}
## R code 4.32
vcov( m4.1 )

## R code 4.33
diag( vcov( m4.1 ) )
cov2cor( vcov( m4.1 ) )

## R code 4.34
library(rethinking)
post <- extract.samples( m4.1 , n=1e4 )
head(post)

## R code 4.35
precis(post)

## R code 4.36
library(MASS)
post <- mvrnorm( n=1e4 , mu=coef(m4.1) , Sigma=vcov(m4.1) )
```

## 4.4. Adding a predictor

```{r}
## R code 4.37
plot( d2$height ~ d2$weight )
```

### 4.4.1. The linear model strategy.

```{r}
## R code 4.38
set.seed(2971)
N <- 100                   # 100 lines
a <- rnorm( N , 178 , 20 )
b <- rnorm( N , 0 , 10 )

## R code 4.39
plot( NULL , xlim=range(d2$weight) , ylim=c(-100,400) ,
    xlab="weight" , ylab="height" )
abline( h=0 , lty=2 )
abline( h=272 , lty=1 , lwd=0.5 )
mtext( "b ~ dnorm(0,10)" )
xbar <- mean(d2$weight)
for ( i in 1:N ) curve( a[i] + b[i]*(x - xbar) ,
    from=min(d2$weight) , to=max(d2$weight) , add=TRUE ,
    col=col.alpha("black",0.2) )

## R code 4.40
b <- rlnorm( 1e4 , 0 , 1 )
dens( b , xlim=c(0,5) , adj=0.1 )

## R code 4.41
set.seed(2971)
N <- 100                   # 100 lines
a <- rnorm( N , 178 , 20 )
b <- rlnorm( N , 0 , 1 )

## R code 4.39
plot( NULL , xlim=range(d2$weight) , ylim=c(-100,400) ,
    xlab="weight" , ylab="height" )
abline( h=0 , lty=2 )
abline( h=272 , lty=1 , lwd=0.5 )
mtext( "b ~ dnorm(0,10)" )
xbar <- mean(d2$weight)
for ( i in 1:N ) curve( a[i] + b[i]*(x - xbar) ,
    from=min(d2$weight) , to=max(d2$weight) , add=TRUE ,
    col=col.alpha("black",0.2) )

## R code 4.40
b <- rlnorm( 1e4 , 0 , 1 )
dens( b , xlim=c(0,5) , adj=0.1 )
```

### 4.4.2. Finding the posterior distribution.

```{r}
## R code 4.42
# load data again, since it's a long way back
library(rethinking)
data(Howell1)
d <- Howell1
d2 <- d[ d$age >= 18 , ]

# define the average weight, x-bar
xbar <- mean(d2$weight)

# fit model
m4.3 <- quap(
    alist(
        height ~ dnorm( mu , sigma ) ,
        mu <- a + b*( weight - xbar ) ,
        a ~ dnorm( 178 , 20 ) ,
        b ~ dlnorm( 0 , 1 ) ,
        sigma ~ dunif( 0 , 50 )
    ) ,
    data=d2 )

## R code 4.43
m4.3b <- quap(
    alist(
        height ~ dnorm( mu , sigma ) ,
        mu <- a + exp(log_b)*( weight - xbar ),
        a ~ dnorm( 178 , 100 ) ,
        log_b ~ dnorm( 0 , 1 ) ,
        sigma ~ dunif( 0 , 50 )
    ) ,
    data=d2 )
```

## 4.7. Practice

### Easy.

#### 4E1. In the model definition below, which line is the likelihood?

yi ~ Normal(u, sigma)

> the likelihood

u ~ Normal(0, 10)

> the prior for u (mean)

sigma ~ Uniform(0, 10)

> the prior for sigma (SD)

#### 4E2. In the model definition just above, how many parameters are in the posterior distribution?

> Two, u and sigma are parameters. yi is observed data.

#### 4E3. Using the model definition above, write down the appropriate form of Bayes’ theorem that includes the proper likelihood and priors.

$\Pr(μ,σ|y)=\prod_iNormal(yi|μ,σ)Normal(μ|0,10)Uniform(σ|0,10)/\int\int\prod_iNormal(hi|μ,σ)Normal(μ|0,10)Uniform(σ|0,10)dμdσ$

> Pr(μ,σ|y)=∏iNormal(yi|μ,σ)Normal(μ|0,10)Uniform(σ|0,10)/∫∫∏iNormal(hi|μ,σ)Normal(μ|0,10)Uniform(σ|0,10)dμdσ

#### 4E4. In the model definition below, which line is the linear model?
yi ~ Normal(u, sigma)

> the likelihood

ui = a + bxi
> the linear model

a ~ Normal(0, 10)

> the prior for a

b ~ Normal(0, 1)

> the prior for b

sigma ~ Uniform(0, 10)

> the prior for sigma

#### 4E5. In the model definition just above, how many parameters are in the posterior distribution?

> Three, a, b, and sigma are parameters.


### Medium.

#### 4M1. For the model definition below, simulate observed heights from the prior (not the posterior).
#### yi ~ Normal(u, sigma)
#### u ~ Normal(0, 10)
#### sigma ~ Uniform(0, 10)

```{r}
sample_mu <- rnorm( 1e4 , 0 , 10 )
sample_sigma <- runif( 1e4 , 0 , 10 )
prior_y<- rnorm( 1e4 , sample_mu , sample_sigma )
dens(prior_y)
```


#### 4M2. Translate the model just above into a quap formula.

```{r}
## R code 4.27
formula4M2 <- alist(
    y ~ dnorm(mu,sigma) ,
    mu ~ dnorm(0,10) ,
    sigma ~ dunif(0,20)
)
```


```{r}
## R code 4.28
data2 <- data.frame(yi=prior_y)
m4M2 <- quap( formula4M2,data = data2,start=list(mu=0,sigma=5))
d2
data2
start
```


#### 4M3. Translate the quap model formula below into a mathematical model definition.

```{r}
flist <- alist(
y ~ dnorm( mu , sigma ),
mu <- a + b*x,
a ~ dnorm( 0 , 50 ),
b ~ dunif( 0 , 10 ),
sigma ~ dunif( 0 , 50 )
)
```

> yi ~ Normal(mu, sigma)

> mu <- a + b*xi

> a ~ Normal(0, 50)

> b ~ Uniform(0, 10)

> sigma ~ Uniform(0, 50)