---
title: "chapter 4-2"
author: "Min-Yao"
date: "2019年4月30日"
output: 
  html_document: 
    keep_md: yes
---

```{r}
## R code 4.42
# load data again, since it's a long way back
library(rethinking)
data(Howell1)
d <- Howell1
d2 <- d[ d$age >= 18 , ]

# define the average weight, x-bar
xbar <- mean(d2$weight)

# fit model
m4.3 <- quap(
    alist(
        height ~ dnorm( mu , sigma ) ,
        mu <- a + b*( weight - xbar ) ,
        a ~ dnorm( 178 , 20 ) ,
        b ~ dlnorm( 0 , 1 ) ,
        sigma ~ dunif( 0 , 50 )
    ) ,
    data=d2 )

## R code 4.43
m4.3b <- quap(
    alist(
        height ~ dnorm( mu , sigma ) ,
        mu <- a + exp(log_b)*( weight - xbar ),
        a ~ dnorm( 178 , 100 ) ,
        log_b ~ dnorm( 0 , 1 ) ,
        sigma ~ dunif( 0 , 50 )
    ) ,
    data=d2 )

## R code 4.44
precis( m4.3 )
```

```{r}
## R code 4.45
round( vcov( m4.3 ) , 3 )
pairs(m4.3)
```

```{r}
## R code 4.46
plot( height ~ weight , data=d2 , col=rangi2 )
post <- extract.samples( m4.3 )
a_map <- mean(post$a)
b_map <- mean(post$b)
curve( a_map + b_map*(x - xbar) , add=TRUE )
```

```{r}
## R code 4.47
post <- extract.samples( m4.3 )
post[1:5,]
```

```{r} 
## R code 4.48
N <- 10
dN <- d2[ 1:N , ]
mN <- quap(
    alist(
        height ~ dnorm( mu , sigma ) ,
        mu <- a + b*( weight - mean(weight) ) ,
        a ~ dnorm( 178 , 20 ) ,
        b ~ dlnorm( 0 , 1 ) ,
        sigma ~ dunif( 0 , 50 )
    ) , data=dN )

## R code 4.49
# extract 20 samples from the posterior
post <- extract.samples( mN , n=20 )

# display raw data and sample size
plot( dN$weight , dN$height ,
    xlim=range(d2$weight) , ylim=range(d2$height) ,
    col=rangi2 , xlab="weight" , ylab="height" )
mtext(concat("N = ",N))

# plot the lines, with transparency
for ( i in 1:20 )
    curve( post$a[i] + post$b[i]*(x-mean(dN$weight)) ,
        col=col.alpha("black",0.3) , add=TRUE )
```

```{r} 
## R code 4.48
N <- 100
dN <- d2[ 1:N , ]
mN <- quap(
    alist(
        height ~ dnorm( mu , sigma ) ,
        mu <- a + b*( weight - mean(weight) ) ,
        a ~ dnorm( 178 , 20 ) ,
        b ~ dlnorm( 0 , 1 ) ,
        sigma ~ dunif( 0 , 50 )
    ) , data=dN )

## R code 4.49
# extract 20 samples from the posterior
post <- extract.samples( mN , n=20 )

# display raw data and sample size
plot( dN$weight , dN$height ,
    xlim=range(d2$weight) , ylim=range(d2$height) ,
    col=rangi2 , xlab="weight" , ylab="height" )
mtext(concat("N = ",N))

# plot the lines, with transparency
for ( i in 1:20 )
    curve( post$a[i] + post$b[i]*(x-mean(dN$weight)) ,
        col=col.alpha("black",0.3) , add=TRUE )
```

```{r} 
## R code 4.48
N <- 300
dN <- d2[ 1:N , ]
mN <- quap(
    alist(
        height ~ dnorm( mu , sigma ) ,
        mu <- a + b*( weight - mean(weight) ) ,
        a ~ dnorm( 178 , 20 ) ,
        b ~ dlnorm( 0 , 1 ) ,
        sigma ~ dunif( 0 , 50 )
    ) , data=dN )

## R code 4.49
# extract 20 samples from the posterior
post <- extract.samples( mN , n=20 )

# display raw data and sample size
plot( dN$weight , dN$height ,
    xlim=range(d2$weight) , ylim=range(d2$height) ,
    col=rangi2 , xlab="weight" , ylab="height" )
mtext(concat("N = ",N))

# plot the lines, with transparency
for ( i in 1:20 )
    curve( post$a[i] + post$b[i]*(x-mean(dN$weight)) ,
        col=col.alpha("black",0.3) , add=TRUE )
```

```{r}
## R code 4.50
post <- extract.samples( m4.3 )
mu_at_50 <- post$a + post$b * ( 50 - xbar )
```

```{r}
## R code 4.51
dens( mu_at_50 , col=rangi2 , lwd=2 , xlab="mu|weight=50" )
```

```{r}
## R code 4.52
HPDI( mu_at_50 , prob=0.89 )
```

```{r}
## R code 4.53
mu <- link( m4.3 )
str(mu)
```

```{r}
## R code 4.54
# define sequence of weights to compute predictions for
# these values will be on the horizontal axis
weight.seq <- seq( from=25 , to=70 , by=1 )

# use link to compute mu
# for each sample from posterior
# and for each weight in weight.seq
mu <- link( m4.3 , data=data.frame(weight=weight.seq) )
str(mu)
```

```{r}
## R code 4.55
# use type="n" to hide raw data
plot( height ~ weight , d2 , type="n" )

# loop over samples and plot each mu value
for ( i in 1:100 )
    points( weight.seq , mu[i,] , pch=16 , col=col.alpha(rangi2,0.1) )
```

```{r}
## R code 4.56
# summarize the distribution of mu
mu.mean <- apply( mu , 2 , mean )
mu.HPDI <- apply( mu , 2 , HPDI , prob=0.89 )
mu.mean
mu.HPDI
```

```{r}
## R code 4.57
# plot raw data
# fading out points to make line and interval more visible
plot( height ~ weight , data=d2 , col=col.alpha(rangi2,0.5) )

# plot the MAP line, aka the mean mu for each weight
lines( weight.seq , mu.mean )

# plot a shaded region for 89% HPDI
shade( mu.HPDI , weight.seq )
```

```{r}
## R code 4.58
post <- extract.samples(m4.3)
mu.link <- function(weight) post$a + post$b*( weight - xbar )
weight.seq <- seq( from=25 , to=70 , by=1 )
mu <- sapply( weight.seq , mu.link )
mu.mean <- apply( mu , 2 , mean )
mu.HPDI <- apply( mu , 2 , HPDI , prob=0.89 )
mu.mean
mu.HPDI
```


```{r}
## R code 4.59
sim.height <- sim( m4.3 , data=list(weight=weight.seq) )
str(sim.height)

## R code 4.60
height.PI <- apply( sim.height , 2 , PI , prob=0.89 )

## R code 4.61
# plot raw data
plot( height ~ weight , d2 , col=col.alpha(rangi2,0.5) )

# draw MAP line
lines( weight.seq , mu.mean )

# draw HPDI region for line
shade( mu.HPDI , weight.seq )

# draw PI region for simulated heights
shade( height.PI , weight.seq )
```
```{r}
## R code 4.59
sim.height <- sim( m4.3 , data=list(weight=weight.seq) )
str(sim.height)

## R code 4.60
height.PI <- apply( sim.height , 2 , PI , prob=0.67 )

## R code 4.61
# plot raw data
plot( height ~ weight , d2 , col=col.alpha(rangi2,0.5) )

# draw MAP line
lines( weight.seq , mu.mean )

# draw HPDI region for line
shade( mu.HPDI , weight.seq )

# draw PI region for simulated heights
shade( height.PI , weight.seq )
```

```{r}
## R code 4.62
sim.height <- sim( m4.3 , data=list(weight=weight.seq) , n=1e4 )
height.PI <- apply( sim.height , 2 , PI , prob=0.89 )
```

```{r}
## R code 4.61
# plot raw data
plot( height ~ weight , d2 , col=col.alpha(rangi2,0.5) )

# draw MAP line
lines( weight.seq , mu.mean )

# draw HPDI region for line
shade( mu.HPDI , weight.seq )

# draw PI region for simulated heights
shade( height.PI , weight.seq )
```

```{r}
## R code 4.63
post <- extract.samples(m4.3)
weight.seq <- 25:70
sim.height <- sapply( weight.seq , function(weight)
    rnorm(
        n=nrow(post) ,
        mean=post$a + post$b*( weight - xbar ) ,
        sd=post$sigma ) )
height.PI <- apply( sim.height , 2 , PI , prob=0.89 )
```

```{r}
## R code 4.64
library(rethinking)
data(Howell1)
d <- Howell1
str(d)

## R code 4.68
plot( height ~ weight , d , col=col.alpha(rangi2,0.5) )
```

```{r}
## R code 4.65
d$weight_s <- ( d$weight - mean(d$weight) )/sd(d$weight)
d$weight_s2 <- d$weight_s^2
m4.5 <- quap(
    alist(
        height ~ dnorm( mu , sigma ) ,
        mu <- a + b1*weight_s + b2*weight_s2 ,
        a ~ dnorm( 178 , 20 ) ,
        b1 ~ dlnorm( 0 , 1 ) ,
        b2 ~ dnorm( 0 , 1 ) ,
        sigma ~ dunif( 0 , 50 )
    ) ,
    data=d )

## R code 4.66
precis( m4.5 )
```

```{r}
## R code 4.67
weight.seq <- seq( from=-2.2 , to=2 , length.out=30 )
pred_dat <- list( weight_s=weight.seq , weight_s2=weight.seq^2 )
mu <- link( m4.5 , data=pred_dat )
mu.mean <- apply( mu , 2 , mean )
mu.PI <- apply( mu , 2 , PI , prob=0.89 )
sim.height <- sim( m4.5 , data=pred_dat )
height.PI <- apply( sim.height , 2 , PI , prob=0.89 )

## R code 4.68
plot( height ~ weight_s , d , col=col.alpha(rangi2,0.5) )
lines( weight.seq , mu.mean )
shade( mu.PI , weight.seq )
shade( height.PI , weight.seq )
```

```{r}
## R code 4.69
d$weight_s3 <- d$weight_s^3
m4.6 <- quap(
    alist(
        height ~ dnorm( mu , sigma ) ,
        mu <- a + b1*weight_s + b2*weight_s2 + b3*weight_s3 ,
        a ~ dnorm( 178 , 20 ) ,
        b1 ~ dlnorm( 0 , 1 ) ,
        b2 ~ dnorm( 0 , 10 ) ,
        b3 ~ dnorm( 0 , 10 ) ,
        sigma ~ dunif( 0 , 50 )
    ) ,
    data=d )

## R code 4.70
plot( height ~ weight_s , d , col=col.alpha(rangi2,0.5) , xaxt="n" )

## R code 4.71
at <- c(-2,-1,0,1,2)
labels <- at*sd(d$weight) + mean(d$weight)
axis( side=1 , at=at , labels=round(labels,1) )
```

```{r}
## R code 4.72
library(rethinking)
data(cherry_blossoms)
d <- cherry_blossoms
precis(d)
```

```{r}
## R code 4.73
d2 <- d[ complete.cases(d$temp) , ] # complete cases on temp
num_knots <- 5
knot_list <- quantile( d2$year , probs=seq(0,1,length.out=num_knots) )
knot_list
```

```{r}
## R code 4.74
library(splines)
B <- bs(d2$year,
    knots=knot_list[-c(1,num_knots)] ,
    degree=3 , intercept=TRUE )

## R code 4.75
plot( NULL , xlim=range(d2$year) , ylim=c(0,1) , xlab="year" , ylab="basis value" )
for ( i in 1:ncol(B) ) lines( d2$year , B[,i] )
```

```{r}
## R code 4.76
m4.7 <- quap(
    alist(
        T ~ dnorm( mu , sigma ) ,
        mu <- a + B %*% w ,
        a ~ dnorm(6,10),
        w ~ dnorm(0,1),
        sigma ~ dexp(1)
    ),
    data=list( T=d2$temp , B=B ) ,
    start=list( w=rep( 0 , ncol(B) ) ) )

## R code 4.77
post <- extract.samples(m4.7)
w <- apply( post$w , 2 , mean )
plot( NULL , xlim=range(d2$year) , ylim=c(-2,2) ,
    xlab="year" , ylab="basis * weight" )
for ( i in 1:ncol(B) ) lines( d2$year , w[i]*B[,i] )
```

```{r}
## R code 4.78
mu <- link( m4.7 )
mu_PI <- apply(mu,2,PI,0.97)
plot( d2$year , d2$temp , col=col.alpha(rangi2,0.3) , pch=16 )
shade( mu_PI , d2$year , col=col.alpha("black",0.5) )

```

## 4.7. Practice

### 4M4, A sample of students is measured for height each year for 3 years. After the third year, you want to fit a linear regression predicting height using year as a predictor. Write down the mathematical model definition for this regression, using any variable names and priors you choose. Be prepared to defend your choice of priors.

> hi ~ Normal(mu, sigma)

> mu <- a + b*xi

> a ~ Normal(163, 30)

> b ~ Normal(2, 1)

> sigma ~ Uniform(0, 20)

> I assume these are college ages students in UC Davis. For the prior a, I chose a normal distribution centered on 163 cm with an SD of 30 cm (because 163 cm is my height). And the prior b, I chose a normal distribution centered on 2 cm/year with an SD of 1 cm/year (because I don't think I will grow any taller). I chose a uniform distribution from 0 cm to 20 cm.


### 4M5, Now suppose I tell you that the average height in the first year was 120 cm and that every student got taller each year. Does this information lead you to change your choice of priors? How?

> Yes! These data are not from college ages students. I changed the prior a, I chose a normal distribution centered on 120 cm with an SD of 30 cm. Also, I changed the prior b, I chose a normal distribution centered on 5 cm/year with an SD of 3 cm/year.

> hi ~ Normal(mu, sigma)

> mu <- a + b*xi

> a ~ Normal(120, 30)

> b ~ Normal(5, 3)

> sigma ~ Uniform(0, 20)

### 4M6, Now suppose I tell you that the variance among heights for students of the same age is never more than 64cm. How does this lead you to revise your priors? 

> Because the variance is the square of SD, if we know that the variance among heights for students of the same age is never more than 64 cm, then SD should be less than 8 cm.

> hi ~ Normal(mu, sigma)

> mu <- a + b*xi

> a ~ Normal(120, 30)

> b ~ Normal(5, 3)

> sigma ~ Uniform(0, 8)

### 4H1, The weights listed below were recorded in the !Kung census, but heights were not recorded for these individuals. Provide predicted heights and 89% intervals (either HPDI or PI) for each of these individuals. That is, fill in the table below, using model-based predictions.

Individual weight expected height 89% interval
1 46.95
2 43.72
3 64.78
4 32.59
5 54.63

```{r}
## change from R code 4.42
# load data again, since it's a long way back
library(rethinking)
data(Howell1)
d <- Howell1

# fit model
m4H1 <- quap(
    alist(
        height ~ dnorm( mu , sigma ) ,
        mu <- a + b* weight ,
        a ~ dnorm( 178 , 20 ) ,
        b ~ dlnorm( 0 , 1 ) ,
        sigma ~ dunif( 0 , 50 )
    ) ,
    data=d )

## change from R code 4.54
# define sequence of weights to compute predictions 
i_weight <- c(46.95, 43.72, 64.78, 32.59, 54.63)

# use link to compute expected height
# for each sample from posterior
expected_height <- link( m4H1 , data=data.frame(weight=i_weight) )
str(expected_height)
```

```{r}
## change from R code 4.56
# summarize the distribution of expected_height
height.mean <- apply( expected_height , 2 , mean )
height.HPDI <- apply( expected_height , 2 , HPDI , prob=0.89 )
height.mean
height.HPDI

new_table = data.frame( 
  individual = 1:5,  
  weight = i_weight,  
  expected_height = height.mean,  
  HPDI89_lower = height.HPDI[1, ],  
  HPDI89_upper = height.HPDI[2, ] 
)
new_table

```

```{r}
## change from R code 4.59
sim.height <- sim( m4H1 , data=list(weight=i_weight) )
str(sim.height)

## change from R code 4.60
sim_height.mean <- apply( sim.height , 2 , mean )
sim_height.HPDI <- apply( sim.height , 2 , HPDI , prob=0.89 )
sim_height.mean
sim_height.HPDI

new_table2 = data.frame( 
  individual = 1:5,  
  weight = i_weight,  
  expected_height = sim_height.mean,  
  HPDI89_lower = sim_height.HPDI[1, ],  
  HPDI89_upper = sim_height.HPDI[2, ] 
)
new_table2
```


### 4H2 Select out all the rows in the Howell1 data with ages below 18 years of age. If you do it right, you should end up with a new data frame with 192 rows in it.

```{r}
## change from R code 4.42
# load data again, since it's a long way back
library(rethinking)
data(Howell1)
d <- Howell1
d4H2 <- d[ d$age < 18 , ]
str(d4H2)
```


#### (a) Fit a linear regression to these data, using quap. Present and interpret the estimates. For every 10 units of increase in weight, how much taller does the model predict a child gets?

```{r}
# fit model
m4H2 <- quap(
    alist(
        height ~ dnorm( mu , sigma ) ,
        mu <- a + b* weight ,
        a ~ dnorm( 178 , 20 ) ,
        b ~ dlnorm( 0 , 1 ) ,
        sigma ~ dunif( 0 , 50 )
    ) ,
    data=d4H2 )

precis(m4H2)
```

> For every 10 units of increase in weight, the model predict a child gets 26.9 cm taller.

#### (b) Plot the raw data, with height on the vertical axis and weight on the horizontal axis. Super-impose the MAP regression line and 89% HPDI for the mean. Also superimpose the 89% HPDI for predicted heights.

```{r}
## change from R code 4.54
# define sequence of weights to compute predictions for
# these values will be on the horizontal axis
weight.seq <- seq( from= min(d4H2$weight) , to= max(d4H2$weight) , by=1 )

# use link to compute mu
# for each sample from posterior
# and for each weight in weight.seq
mu <- link( m4H2 , data=data.frame(weight=weight.seq) )
str(mu)

## change from R code 4.56
# summarize the distribution of mu
mu.mean <- apply( mu , 2 , mean )
mu.HPDI <- apply( mu , 2 , HPDI , prob=0.89 )
mu.mean
mu.HPDI

## change from R code 4.61
# plot raw data
plot( height ~ weight , d4H2 , col=col.alpha(rangi2,0.5) )

# draw MAP line
lines( weight.seq , mu.mean )

# draw HPDI region for line
shade( mu.HPDI , weight.seq )



## R code 4.59
sim.height <- sim( m4H2 , data=list(weight=weight.seq) )
str(sim.height)

## R code 4.60
sim.height.HPDI <- apply( sim.height , 2 , HPDI , prob=0.89 )

# draw PI region for simulated heights
shade( sim.height.HPDI , weight.seq )

```


#### (c) What aspects of the model fit concern you? Describe the kinds of assumptions you would change, if any, to improve the model. You don’t have to write any new code. Just explain what the model appears to be doing a bad job of, and what you hypothesize would be a better model.

> I think a parabolic (second order) polynomial would be a better model.

```{r}
## R code 4.65
d4H2$weight_s <- ( d4H2$weight - mean(d4H2$weight) )/sd(d4H2$weight)
d4H2$weight_s2 <- d4H2$weight_s^2
m4H2c <- quap(
    alist(
        height ~ dnorm( mu , sigma ) ,
        mu <- a + b1*weight_s + b2*weight_s2 ,
        a ~ dnorm( 178 , 20 ) ,
        b1 ~ dlnorm( 0 , 1 ) ,
        b2 ~ dnorm( 0 , 1 ) ,
        sigma ~ dunif( 0 , 50 )
    ) ,
    data=d4H2 )

precis( m4H2c )

```


```{r}
## change from R code 4.67
weight.seq <- seq( from=-2.2 , to=2 , length.out=30 )
pred_dat <- list( weight_s=weight.seq , weight_s2=weight.seq^2 )
mu <- link( m4H2c , data=pred_dat )

mu.mean <- apply( mu , 2 , mean )
mu.PI <- apply( mu , 2 , PI , prob=0.89 )

sim.height <- sim( m4H2c , data=pred_dat )
height.PI <- apply( sim.height , 2 , PI , prob=0.89 )

## change from R code 4.68
plot( height ~ weight_s , d4H2 , col=col.alpha(rangi2,0.5) )
lines( weight.seq , mu.mean )
shade( mu.PI , weight.seq )
shade( height.PI , weight.seq )
```

