---
title: "chapter14-7_phylogenetic_regression"
author: "Min-Yao"
date: "2020/7/6"
output: 
  html_document: 
    keep_md: yes
---

## You want to rest the hypothesis that the range size of carnivores is influenced by their body size (larger carnivores roam further).  You want to be sure that if you do see a relationship that is is not just due to shared evolutionary history.  That is you want to do a phylogenetic regression.

# 1) Load the data as follows

```{r}
library(rethinking)
library(ape)
library(ade4)
library(stringr)
```

```{r}
data("carni70")
```

```{r}
tre <- read.tree(text=carni70$tre)
d <- carni70$tab
str(d)
summary(d)
```

```{r}
plot( ladderize(tre) , type="fan" , font=1 , no.margin=TRUE ,
    label.offset=1 , cex=0.5 )
```

# 2) Note that the species name given in "rownames(d)" and in "tre$tip.label" do not match.  Fix that (e.g. using str_replace() )

```{r}
head(rownames(d))
head(tre$tip.label)
```
```{r}
rownames(d) = str_replace(rownames(d), "_", ".")

head(rownames(d))
head(tre$tip.label)
```

# 3) Should either range or size be transformed?

```{r}
str(d)
summary(d)
head(d)
```
```{r}
dens(d$size)
dens(standardize(log(d$size)))
```
```{r}
dens(d$range)
dens(standardize(d$range))
dens(standardize(log(d$range)))
```

# 4) Does size predict range?  Evaluate this question without and with accounting for phylogenetic relationships.

```{r}
plot( d$size , d$range , xlab="size" , ylab="range" )
plot( standardize(log(d$size)) , standardize(d$range) , xlab="log_size" , ylab="log_range" )
plot( standardize(log(d$size)) , standardize(log(d$range)) , xlab="log_size" , ylab="log_range" )
```
```{r}
## R code 14.47
d$name <- as.character(rownames(d))
spp_obs <- d$name
```

```{r}
## modified from R code 14.48
dat_list <- list(
    N_spp = nrow(d),
    S = standardize(log(d$size)),
    R = standardize(log(d$range)),
    Imat = diag( nrow(d) )
)

str(dat_list)
```
> Evaluate this question without accounting for phylogenetic relationships

> Using an identity matrix (no phylogeny)

```{r}
m14_7_without <- ulam(
    alist(
        R ~ multi_normal( mu , SIGMA ),
        mu <- a + bS*S ,
        matrix[N_spp,N_spp]: SIGMA <- Imat * sigma_sq,
        a ~ normal( 0 , 1 ),
        bS ~ normal( 0 , 0.5 ),
        sigma_sq ~ exponential( 1 )
    ), data=dat_list , chains=4 , cores=4 )
precis( m14_7_without )
plot(precis( m14_7_without ))
```

> Evaluate this question with accounting for phylogenetic relationships

> Using the Brownian motion model

```{r}
## modified from R code 14.49

Rbm <- corBrownian( phy=tre , form = ~tre$tip.label)
V <- vcv(Rbm)
#error: No covariate specified, species will be taken as ordered in the data frame. To avoid this message, specify a covariate containing the species names with the 'form' argument.
Dmat <- cophenetic( tre )
plot( Dmat , V , xlab="phylogenetic distance" , ylab="covariance" )
image(V)
image(Dmat)
```

```{r}
## modified from R code 14.50
# put species in right order
dat_list$V <- V[ spp_obs , spp_obs ] #subscript out of bounds?

# convert to correlation matrix
dat_list$CoR <- dat_list$V / max(V)

# Brownian motion model
m14_7_with <- ulam(
    alist(
        R ~ multi_normal( mu , SIGMA ),
        mu <- a + bS*S ,
        matrix[N_spp,N_spp]: SIGMA <- CoR * sigma_sq,
        a ~ normal( 0 , 1 ),
        bS ~ normal( 0 , 0.5 ),
        sigma_sq ~ exponential( 1 )
    ), data=dat_list , chains=4 , cores=4 )
precis( m14_7_with )
plot(precis( m14_7_with ))
```

> Using the Ornstein–Uhlenbeck process (or OU process)

> use the OU process kernel, which is known more generally as the L1 norm, which ulam() provides as cov_GPL1().

```{r}
## R code 14.51
# add scaled and reordered distance matrix
dat_list$Dmat <- Dmat[ spp_obs , spp_obs ] / max(Dmat)

m14_7_with_OU <- ulam(
    alist(
        R ~ multi_normal( mu , SIGMA ),
        mu <- a + bS*S ,
        matrix[N_spp,N_spp]: SIGMA <- cov_GPL1( Dmat , etasq , rhosq , 0.01 ),
        a ~ normal(0,1),
        bS ~ normal(0,0.5),
        etasq ~ half_normal(1,0.25),
        rhosq ~ half_normal(3,0.25)
    ), data=dat_list , chains=4 , cores=4 )
precis( m14_7_with_OU )
plot(precis( m14_7_with_OU ))
```

# 5) Discuss your findings.

> Using an identity matrix (no phylogeny), body size influences the range size of carnivores positively 

> Using the Brownian motion model, body size influences the range size of carnivores even more positively.

> Using the Ornstein–Uhlenbeck process (or OU process), body size influences the range size of carnivores even more positively.

```{r}
precis( m14_7_without )
plot(precis( m14_7_without ))
precis( m14_7_with )
plot(precis( m14_7_with ))
precis( m14_7_with_OU )
plot(precis( m14_7_with_OU ))
```






