---
title: "chapter16-1"
author: "Min-Yao"
date: "2020/8/2"
output: 
  html_document: 
    keep_md: yes
---

# Generalized Linear Madness

## 16.1. Geometric people
### 16.1.1. The scientific model.
### 16.1.2. The statistical model.

```{r}
## R code 16.1
library(rethinking)
data(Howell1)
d <- Howell1

# scale observed variables
d$w <- d$weight / mean(d$weight)
d$h <- d$height / mean(d$height)

## R code 16.2
m16.1 <- ulam(
    alist(
        w ~ dlnorm( mu , sigma ),
        exp(mu) <- 3.141593 * k * p^2 * h^3,
        p ~ beta( 2 , 18 ),
        k ~ exponential( 0.5 ),
        sigma ~ exponential( 1 )
    ), data=d , chains=4 , cores=4 )
precis(m16.1)
pairs(m16.1)
```

```{r}
## R code 16.3
h_seq <- seq( from=0 , to=max(d$h) , length.out=30 )
w_sim <- sim( m16.1 , data=list(h=h_seq) )
mu_mean <- apply( w_sim , 2 , mean )
w_CI <- apply( w_sim , 2 , PI )
plot( d$h , d$w , xlim=c(0,max(d$h)) , ylim=c(0,max(d$w)) , col=rangi2 ,
    lwd=2 , xlab="height (scaled)" , ylab="weight (scaled)" )
lines( h_seq , mu_mean )
shade( w_CI , h_seq )
```

### 16.1.3. GLM in disguise.
## 16.2. Hidden minds and observed behavior

```{r}
## R code 16.4
library(rethinking)
data(Boxes)
precis(Boxes)

## R code 16.5
table( Boxes$y ) / length( Boxes$y )
```

### 16.2.1. The scientific model.

```{r}
## R code 16.6
set.seed(7)
N <- 30 # number of children

# half are random
# sample from 1,2,3 at random for each
y1 <- sample( 1:3 , size=N/2 , replace=TRUE )

# half follow majority
y2 <- rep( 2 , N/2 )

# combine and shuffle y1 and y2
y <- sample( c(y1,y2) )

# count the 2s
sum(y==2)/N
```

### 16.2.2. The statistical model.
### 16.2.3. Coding the statistical model.

```{r}
## R code 16.7
data(Boxes_model)
cat(Boxes_model)

## R code 16.8
# prep data
dat_list <- list(
    N = nrow(Boxes),
    y = Boxes$y,
    majority_first = Boxes$majority_first )

# run the sampler
m16.2 <- stan( model_code=Boxes_model , data=dat_list , chains=3 , cores=3 )

# show marginal posterior for p
p_labels <- c("1 Majority","2 Minority","3 Maverick","4 Random","5 Follow First")
plot( precis(m16.2,2) , labels=p_labels )
```

### 16.2.4. State space models.

#### 16M1. Modify the cylinder height model, m16.1, so that the exponent 3 on height is instead a free parameter. Do you recover the value of 3 or not? Plot the posterior predictions for the new model. How do they differ from those of m16.1?

```{r}
## modify from R code 16.2
m16M1 <- ulam(
    alist(
        w ~ dlnorm( mu , sigma ),
        exp(mu) <- 3.141593 * k * p^2 * h^e,
        p ~ beta( 2 , 18 ),
        k ~ exponential( 0.5 ),
        e ~ normal(3,5),
        sigma ~ exponential( 1 )
    ), data=d , chains=4 , cores=4 )
precis(m16M1)
pairs(m16M1)
```

> No, it doen't recover the value of 3. It's about 2.33.

> Plot the posterior predictions for the new model.

```{r}
## modify from R code 16.3
h_seq <- seq( from=0 , to=max(d$h) , length.out=30 )
w_sim <- sim( m16M1 , data=list(h=h_seq) )
mu_mean <- apply( w_sim , 2 , mean )
w_CI <- apply( w_sim , 2 , PI )
plot( d$h , d$w , xlim=c(0,max(d$h)) , ylim=c(0,max(d$w)) , col=rangi2 ,
    lwd=2 , xlab="height (scaled)" , ylab="weight (scaled)" )
lines( h_seq , mu_mean )
shade( w_CI , h_seq )
```


#### 16M2. Conduct a prior predictive simulation for the cylinder height model. Begin with the priors in the chapter. Do these produce reasonable prior height distributions? If not, which modifications do you suggest?

```{r}
## R code 16.1
library(rethinking)
data(Howell1)
d <- Howell1
str(d)
## modify from R code 16.10
N <- 1000
p <- rbeta( N , 2 , 18 )
k <- rexp(N, 0.5)
```

```{r}
## modify from R code 4.39
plot( NULL , xlim=range(d$weight) , ylim=c(-100,400) , xlab="weight" , ylab="height" )
abline( h=0 , lty=2 )
abline( h=272 , lty=1 , lwd=0.5 )
for ( i in 1:N ) curve( (3.141593*k[i]*p[i]^2*x^3), from=min(d$weight) , to=max(d$weight) , add=TRUE , col=col.alpha("black",0.2) )

dens( p , lwd=2 )
dens( x , lwd=2 )
```
