---
title: "chapter 11-3 and chapter 12-1"
author: "Min-Yao"
date: "2019/11/21"
output: 
  html_document: 
    keep_md: yes
---

## 11.3. Censoring and survival

```{r}
library(rethinking)
## R code 11.62
N <- 2
x <- replicate( 1e5 , min(runif(N,1,100)) )
dens(x)

## R code 11.63
N <- 10
M <- 2
x <- replicate( 1e5 , sort(runif(N,1,100))[M] )
dens(x)
```

### 11.3.1. Simulated cats.
### 11.3.2. Actual cats.

```{r}
## R code 11.64
library(rethinking)
data(AustinCats)
d <- AustinCats

d$adopt <- ifelse( d$out_event=="Adoption" , 1L , 0L )
dat <- list(
    days_to_event = as.numeric( d$days_to_event ),
    color_id = ifelse( d$color=="Black" , 1L , 2L ) ,
    adopted = d$adopt
)

m11.14 <- ulam(
    alist(
        days_to_event|adopted==1 ~ exponential( lambda ),
        days_to_event|adopted==0 ~ custom(exponential_lccdf( !Y | lambda )),
        lambda <- 1.0/mu,
        log(mu) <- a[color_id],
        a[color_id] ~ normal(0,1)
    ), data=dat , chains=4 , cores=4 )

precis( m11.14 , 2 )

## R code 11.65
post <- extract.samples( m11.14 )
post$D <- exp(post$a)
precis( post , 2 )

### R code 11.66
#model{
#    vector[22356] lambda;
#    b ~ normal( 0 , 1 );
#    a ~ normal( 0 , 1 );
#    for ( i in 1:22356 ) {
#        lambda[i] = a + b * black[i];
#        lambda[i] = exp(lambda[i]);
#    }
#    for ( i in 1:22356 )
#        if ( adopted[i] == 0 ) target += exponential_lccdf(days_to_event[i] | lambda[i]);
#    for ( i in 1:22356 )
#        if ( adopted[i] == 1 ) days_to_event[i] ~ exponential( lambda[i] );
#}

### R code 11.67
#    for ( i in 1:22356 ) {
#        if ( adopted[i] == 0 ) target += -lambda[i]*days_to_event[i];
#        if ( adopted[i] == 1 ) target += log(lambda[i]) - lambda[i]*days_to_event[i];
#    }
```

## 11.4. Summary

### 10H3 (more binomial practice, so relevant to previous weeks, but good for us)
The data contained in library(MASS);data(eagles) are records of salmon pirating attempts by Bald Eagles in Washington State. See ?eagles for details. While one eagle feeds, sometimes another will swoop in and try to steal the salmon from it. Call the feeding eagle the “victim” and the thief the “pirate.” Use the available data to build a binomial GLM of successful pirating attempts.

(a) Consider the following model:
y i ~ Binomial(n i; p i)
log pi/(1 - pi) = a + bP Pi + bV Vi + bA Ai
a ~ Normal (0 ; 10)
bP ~ Normal (0 ; 5)
bV ~ Normal (0 ; 5)
bA ~ Normal (0 ; 5)
where y is the number of successful attempts, n is the total number of attempts, P is a dummy variable indicating whether or not the pirate had large body size, V is a dummy variable indicating whether or not the victim had large body size, and fi nally A is a dummy variable indicating whether or not the pirate was an adult. Fit the model above to the eagles data, using both map and map2stan. Is the quadratic approximation okay?

```{r}
#?eagles
library(MASS)
data(eagles)
d <- eagles
str(d)
head(d)
summary(d)
```

```{r}
d$p_is_l <- ifelse(d$P=='L', 1L, 0L) # P: Size of pirating eagle (L = large, S = small).
d$p_is_a <- ifelse(d$A=='A', 1L, 0L) # A: Age of pirating eagle (I = immature, A = adult).
d$v_is_l <- ifelse(d$V=='L', 1L, 0L) # V: Size of victim eagle (L = large, S = small).

d$prob <- d$y/d$n # y: Number of successful attempts. n: Total number of attempts.

# prior trimmed data list
dat_list <- list(
    p_is_l = d$p_is_l,
    p_is_a = d$p_is_a,
    v_is_l = d$v_is_l,
    y = d$y,
    n = d$n)

str(dat_list)
head(dat_list)
summary(dat_list)
```

```{r}
m10H3.a <- ulam(
    alist(
        y ~ dbinom(n, p),
        logit(p) <- a + bp*p_is_l + bv*v_is_l + ba*p_is_a,
        a ~ dnorm(0, 10),
        c(bp, bv, ba) ~ dnorm(0, 5)), 
    data=dat_list, chains=4, cores = 4, log_lik=TRUE)
precis(m10H3.a)

```


(b) Now interpret the estimates. If the quadratic approximation turned out okay, then it’s okay to use the map estimates. Otherwise stick to map2stan estimates. Then plot the posterior predictions. Compute and display both (1) the predicted probability of success and its 89% interval for each row (i) in the data, as well as (2) the predicted success count and its 89% interval. What different information does each type of posterior prediction provide?

```{r}
plot(precis(m10H3.a))
postcheck(m10H3.a, prob=0.89)
pairs(m10H3.a)
```

```{r}
p <- link(m10H3.a)
y <- sim(m10H3.a)

p.mean <- apply(X = p, MARGIN = 2, FUN = mean)
p.PI <- apply(X = p, MARGIN = 2, FUN = PI)
y.mean <- apply(X = y, MARGIN = 2, FUN = mean)
y.PI <- apply(X = y, MARGIN = 2, FUN = PI)

# plot the model predictions for `p` vs. the actual proportion of successes for each case
d$success.proportion <- d$y / d$n
plot(d$success.proportion, col=rangi2, ylab="successful proportion", xlab="case", xaxt="n", xlim=c(0.75,8.25) , ylim = c(0, 1), pch=16)
axis(1, at=1:8, labels=c( "LAL","LAS","LIL","LIS","SAL","SAS","SIL","SIS" ))
points( 1:8 , p.mean )
for ( i in 1:8 ) lines( c(i, i), p.PI[,i] )
```


(c) Now try to improve the model. Consider an interaction between the pirate’s size and age (immature or adult). Compare this model to the previous one, using WAIC. Interpret.

```{r}
m10H3.c <- ulam(
  alist(
    y ~ dbinom(n, p),
    logit(p) <- a + bp*p_is_l + bv*v_is_l + ba*p_is_a + bpa*p_is_l*p_is_a,
    a ~ dnorm(0, 10),
    c(bp, bv, ba, bpa) ~ dnorm(0, 5)), 
  data=dat_list, chains=4, cores = 4, log_lik=TRUE)
precis(m10H3.c)
pairs(m10H3.c)
compare(m10H3.a, m10H3.c)
postcheck(m10H3.c, prob=0.89)
```


> the model with interaction between the pirate’s size and age looks like better (smaller WAIC and reasonable set of parameters)

# 12 Monsters and Mixtures

## 12.1. Over-dispersed outcomes

### 12.1.1. Beta-binomial.

```{r}
## R code 12.1
pbar <- 0.5
theta <- 5
curve( dbeta2(x,pbar,theta) , from=0 , to=1 ,
    xlab="probability" , ylab="Density" )

pbar <- 0.5
theta <- 2
curve( dbeta2(x,pbar,theta) , from=0 , to=1 ,
    xlab="probability" , ylab="Density" )

pbar <- 0.5
theta <- 1
curve( dbeta2(x,pbar,theta) , from=0 , to=1 ,
    xlab="probability" , ylab="Density" )
```


```{r}
## R code 12.2
library(rethinking)
data(UCBadmit)
d <- UCBadmit
d$gid <- ifelse( d$applicant.gender=="male" , 1L , 2L )
dat <- list( A=d$admit , N=d$applications , gid=d$gid )
m12.1 <- ulam(
    alist(
        A ~ dbetabinom( N , pbar , theta ),
        logit(pbar) <- a[gid],
        a[gid] ~ dnorm( 0 , 1.5 ),
        theta ~ dexp(1)
    ), data=dat , chains=4 )

## R code 12.3
post <- extract.samples( m12.1 )
post$da <- post$a[,1] - post$a[,2]
precis( post , depth=2 )
```

```{r}
## R code 12.4
gid <- 2
# draw posterior mean beta distribution
curve( dbeta2(x,mean(logistic(post$a[,gid])),mean(post$theta)) , from=0 , to=1 ,
    ylab="Density" , xlab="probability admit", ylim=c(0,3) , lwd=2 )

# draw 50 beta distributions sampled from posterior
for ( i in 1:50 ) {
    p <- logistic( post$a[i,gid] )
    theta <- post$theta[i]
    curve( dbeta2(x,p,theta) , add=TRUE , col=col.alpha("black",0.2) )
}
mtext( "distribution of female admission rates" )

## R code 12.5
postcheck( m12.1 )
```

### 12.1.2. Negative-binomial or gamma-Poisson.

```{r}
## R code 12.6
library(rethinking)
data(Kline)
d <- Kline
d$P <- standardize( log(d$population) )
d$contact_id <- ifelse( d$contact=="high" , 2L , 1L )

dat2 <- list(
    T = d$total_tools,
    P = d$population,
    cid = d$contact_id )

m12.3 <- ulam(
    alist(
        T ~ dgampois( lambda , phi ),
        lambda <- exp(a[cid])*P^b[cid] / g,
        a[cid] ~ dnorm(1,1),
        b[cid] ~ dexp(1),
        g ~ dexp(1),
        phi ~ dexp(1)
    ), data=dat2 , chains=4 , log_lik=TRUE )
```

### 12.1.3. Over-dispersion, entropy, and information criteria.

## 12.2. Zero-inflated outcomes

### 12.2.1. Example: Zero-inflated Poisson.

```{r}
## R code 12.7
# define parameters
prob_drink <- 0.2 # 20% of days
rate_work <- 1    # average 1 manuscript per day

# sample one year of production
N <- 365

# simulate days monks drink
set.seed(365)
drink <- rbinom( N , 1 , prob_drink )

# simulate manuscripts completed
y <- (1-drink)*rpois( N , rate_work )

## R code 12.8
simplehist( y , xlab="manuscripts completed" , lwd=4 )
zeros_drink <- sum(drink)
zeros_work <- sum(y==0 & drink==0)
zeros_total <- sum(y==0)
lines( c(0,0) , c(zeros_work,zeros_total) , lwd=4 , col=rangi2 )
```

```{r}
## R code 12.9
m12.4 <- ulam(
    alist(
        y ~ dzipois( p , lambda ),
        logit(p) <- ap,
        log(lambda) <- al,
        ap ~ dnorm( -1.5 , 1 ),
        al ~ dnorm( 1 , 0.5 )
    ) , data=list(y=as.integer(y)) , chains=4 )
precis( m12.4 )

## R code 12.10
inv_logit(-1.28) # probability drink
exp(0.01)       # rate finish manuscripts, when not drinking

## R code 12.11
m12.4_alt <- ulam(
    alist(
        y|y>0 ~ custom( log1m(p) + poisson_lpmf(y|lambda) ),
        y|y==0 ~ custom( log_mix( p , 0 , poisson_lpmf(0|lambda) ) ),
        logit(p) <- ap,
        log(lambda) <- al,
        ap ~ dnorm(-1.5,1),
        al ~ dnorm(1,0.5)
    ) , data=list(y=as.integer(y)) , chains=4 )
```

## 12.6. Practice

### 11E3. When count data are zero-inflated, using a model that ignores zero-inflation will tend to induce which kind of inferential error?

> It will underestimate true value of "lambda" for Poisson regression or parameter "p" for Binomial regression, as it would try to explain large number of zeros with small values of lambda. Using an example from the chapter:

```{r}
## R code 12.7
# define parameters
prob_drink <- 0.2 # 20% of days
rate_work <- 1    # average 1 manuscript per day

# sample one year of production
N <- 365

# simulate days monks drink
set.seed(365)
drink <- rbinom( N , 1 , prob_drink )

# simulate manuscripts completed
y <- (1-drink)*rpois( N , rate_work )

## fit zero-inflated Poisson 
m12.4 <- ulam(
    alist(
        y ~ dzipois( p , lambda ),
        logit(p) <- ap,
        log(lambda) <- al,
        ap ~ dnorm( -1.5 , 1 ),
        al ~ dnorm( 1 , 0.5 )
    ) , data=list(y=as.integer(y)) , chains=4 , log_lik=TRUE)
precis( m12.4 )

## fit ordinary Poisson
m12.4_2 <- ulam(
    alist(
        y ~ dpois(lambda),
        log(lambda) <- al,
        al ~ dnorm( 1 , 0.5 )
    ) , data=list(y=as.integer(y)) , chains=4 , log_lik=TRUE)
precis( m12.4_2 )

compare( m12.4, m12.4_2)
```


### 11E4. Over-dispersion is common in count data. Give an example of a natural process that might produce over-dispersed counts. Can you also give an example of a process that might produce underdispersed counts?

> One of possible artificial examples of under-dispersed variable is a Binomial varibale with all observations smaller than a threshold substituted with the threshold itself. For example, some tomato cultivars are susceptible to Cuscuta and some cultivars are resistant to Cuscuta. If we overexpress a R protein in susceptible clutivars, it will make susceptible plants become resistant. However, if the R gene expression is not high enough in some individual plants, then it won't be able to triger the resistant response. So, these individual plants will still be classified as susceptible.


### 11M3. (optional).  Can you modify the derivation of the zero-inflated Poisson distribution (ZIPoisson) from the chapter to construct a zero-inflated binomial distribution?

> just change Poisson likelihood to Binomial 

 p - probability of zero
 n - number of trials
 L - probability of completed manuscripts with some average rate 
 $Pr(0|p,L) = Pr(drink|p) + Pr(work|p)* Pr(0|L) = p + (1-p) exp(-L)$
 
 yi ~ ZIPoisson(pi; Li)


### 11H6. 
#### The data in data(Fish) are records of visits to a national park. See ?Fish for details. The question of interest is how many fish an average visitor takes per hour, when fishing. The problemis that not everyone tried to fish, so the fish_caught numbers are zero-inflated. As with the monks example in the chapter, there is a process that determines who is fishing (working) and another process that determines fish per hour (manuscripts per day), conditional on fishing (working). We want to model both. Otherwise we’ll end up with an underestimate of rate of fish extraction from the park. You will model these data using zero-inflated Poisson GLMs. Predict fish_caught as a function of any of the other variables you think are relevant. One thing you must do, however, is use a proper Poisson off set/exposure in the Poisson portion of the zero-inflated model. Then usethe hours variable to construct the off set. This will adjust the model for the diff ering amount of time individuals spent in the park.

```{r}
library(rethinking)
data("Fish")
d <- Fish
str(d)
head(d)
summary(d)
#?Fish
#fish_caught : Number of fish caught during visit
#livebait : Whether or not group used livebait to fish
#camper : Whether or not group had a camper
#persons : Number of adults in group
#child : Number of children in group
#hours : Number of hours group spent in park

d$persons_std <- (d$persons - mean(d$persons)) / sd(d$persons)
d$hours_std <- (d$hours - mean(d$hours)) / sd(d$hours)
d$loghours <- log(d$hours)

str(d)
head(d)
summary(d)
```

```{r}
m11H6.1 <- ulam(
  alist(
    fish_caught ~ dzipois(p, lambda),
    logit(p) <-  ap + bp_c*camper + bp_p*persons,
    log(lambda) <- loghours + al + bl_lb*livebait + bl_c*camper + bl_p*persons_std,
    ap ~ dnorm(0,10),
    al ~ dnorm(0,10),
    c(bp_c, bp_p) ~ dnorm(0,2),
    c(bl_lb, bl_c, bl_p) ~ dnorm(0,2)),
  data=d, chains=4 , cores=4, log_lik=TRUE)
precis(m11H6.1)
postcheck(m11H6.1)
```

```{r}
m11H6.0 <- ulam(
  alist(
    fish_caught ~ dzipois(p, lambda),
    logit(p) <-  ap,
    log(lambda) <- loghours + al ,
    ap ~ dnorm(0,10),
    al ~ dnorm(0,10)),
  data=d, chains=4 , cores=4, log_lik=TRUE)
precis(m11H6.0)
postcheck(m11H6.0)
```

```{r}
compare(m11H6.0, m11H6.1)
```


### 11H1. 
#### In 2014, a paper was published that was entitled “Female hurricanes are deadlier than male hurricanes.” 185 As the title suggests, the paper claimed that hurricanes with female names have caused greater loss of life, and the explanation given is that people unconsciously rate female hurricanes as less dangerous and so are less likely to evacuate. Statisticians severely criticized the paper after publication. Here, you’ll explore the complete data used in the paper and consider the hypothesis that hurricanes with female names are deadlier. Load the data with:

```{r}
library(rethinking)
data(Hurricanes)
```


#### Acquaint yourself with the columns by inspecting the help ?Hurricanes. In this problem, you’ll focus on predicting deaths using femininity of each hurricane’s name. Fit and interpret the simplest possible model, a Poisson model of deaths using femininity as a predictor. You can use map or map2stan. Compare the model to an intercept-only Poisson model of deaths. How strong is the association between femininity of name and deaths? Which storms does the model fit (retrodict) well? Which storms does it fit poorly?

```{r}
library(rethinking)
data(Hurricanes)
d <- Hurricanes
str(d)
head(d)
summary(d)
```

```{r}
# prior trimmed data list
dat_list <- list(
    femininity = d$femininity,
    deaths = d$deaths)

# fit Poisson regression using `femininity` as a predictor
m11H1.1 <- ulam(
  alist(
    deaths ~ dpois( lambda ),
    log(lambda) <- a + bf*femininity,
    a ~ dnorm(0, 10),
    bf ~ dnorm(0, 10)),
  data = dat_list, chains=4 , cores=4, log_lik=TRUE)

precis(m11H1.1)
```

```{r}
# fit Poisson regression using intercept only
m11H1.2 <- ulam(
  alist(
    deaths ~ dpois( lambda ),
    log(lambda) <- alpha,
    alpha ~ dnorm(0, 10)),
  data = dat_list, chains=4 , cores=4, log_lik=TRUE)

# compare models
compare(m11H1.1, m11H1.2)
```

```{r}
# plot posterior predictions
prediction.data <- list(femininity = seq(from = 1, to = 11, length.out = 30))

# simulate lambda values
lambda <- link(fit = m11H1.1, data = prediction.data)
lambda.mean <- apply(X = lambda, MARGIN = 2, FUN = mean)
lambda.PI <- apply(X = lambda, MARGIN = 2, FUN = PI)

# simulate Poisson predictions
simulated.predictions <- sim(fit = m11H1.1, data = prediction.data)
simulated.predictions.PI <- apply(X = simulated.predictions, MARGIN = 2, FUN = PI)

plot( d$femininity, d$deaths, pch=16, col=rangi2, xlab="femininity", ylab="deaths" )
lines(x = prediction.data$femininity, y = lambda.mean)
shade(object = lambda.PI, lim = prediction.data$femininity)
lines(x = prediction.data$femininity, y = simulated.predictions.PI[1,], lty = 2)
lines(x = prediction.data$femininity, y = simulated.predictions.PI[2,], lty = 2)
```

> Summary: Intuitively, there is some hidden variable that better explains deaths.

### 11H2. 
#### Counts are nearly always over-dispersed relative to Poisson. So fit a gamma-Poisson (aka negative-binomial) model to predict deaths using femininity. Show that the over-dispersed model no longer shows as precise a positive association between femininity and deaths, with an 89% interval that overlaps zero. Can you explain why the association diminished in strength?

```{r}
# fit model
model.11H2 <- ulam(
  alist(
    deaths ~ dgampois( lambda , phi ),
    log(lambda) <- a + bf*femininity,
    a ~ dnorm(0, 10),
    bf ~ dnorm(0, 10),
    phi ~ dexp(1)
  ),
  data=dat_list,chains=4 , cores=4, log_lik=TRUE)

# inspect estimates
precis(model.11H2)
```

> This may be because a much larger range of values of (alpha, beta_femininity) can produce through the Gamma distribution.
