---
title: "chapter 8-1"
author: "Min-Yao"
date: "2019年7月30日"
output: 
  html_document: 
    keep_md: yes
---

# 8 Conditional Manatees

## 8.1. Building an interaction

### 8.1.1. Making two models.

```{r}
## R code 8.1
library(rethinking)
data(rugged)
d <- rugged

# make log version of outcome
d$log_gdp <- log( d$rgdppc_2000 )

# extract countries with GDP data
dd <- d[ complete.cases(d$rgdppc_2000) , ]

# rescale variables
dd$log_gdp_std <- dd$log_gdp / mean(dd$log_gdp)
dd$rugged_std <- dd$rugged / max(dd$rugged)

# split countries into Africa and not-Africa
d.A1 <- dd[ dd$cont_africa==1 , ] # Africa
d.A0 <- dd[ dd$cont_africa==0 , ] # not Africa

## R code 8.2
m8.1 <- quap(
    alist(
        log_gdp_std ~ dnorm( mu , sigma ) ,
        mu <- a + b*( rugged_std - 0.215 ) ,
        a ~ dnorm( 1 , 1 ) ,
        b ~ dnorm( 0 , 1 ) ,
        sigma ~ dexp( 1 )
    ) , data=d.A1 )

## R code 8.3
set.seed(7)
prior <- extract.prior( m8.1 )

# set up the plot dimensions
plot( NULL , xlim=c(0,1) , ylim=c(0.5,1.5) ,
    xlab="ruggedness" , ylab="log GDP" )
abline( h=min(dd$log_gdp_std) , lty=2 )
abline( h=max(dd$log_gdp_std) , lty=2 )

# draw 50 lines from the prior
rugged_seq <- seq( from=-0.1 , to=1.1 , length.out=30 )
mu <- link( m8.1 , post=prior , data=data.frame(rugged_std=rugged_seq) )
for ( i in 1:50 ) lines( rugged_seq , mu[i,] , col=col.alpha("black",0.3) )
```

```{r}
## R code 8.4
sum( abs(prior$b) > 0.6 ) / length(prior$bR)

## R code 8.5
m8.1 <- quap(
    alist(
        log_gdp_std ~ dnorm( mu , sigma ) ,
        mu <- a + b*( rugged_std - 0.215 ) ,
        a ~ dnorm( 1 , 0.1 ) ,
        b ~ dnorm( 0 , 0.3 ) ,
        sigma ~ dexp(1)
    ) , data=d.A1 )

## R code 8.3
set.seed(7)
prior <- extract.prior( m8.1 )

# set up the plot dimensions
plot( NULL , xlim=c(0,1) , ylim=c(0.5,1.5) ,
    xlab="ruggedness" , ylab="log GDP" )
abline( h=min(dd$log_gdp_std) , lty=2 )
abline( h=max(dd$log_gdp_std) , lty=2 )

# draw 50 lines from the prior
rugged_seq <- seq( from=-0.1 , to=1.1 , length.out=30 )
mu <- link( m8.1 , post=prior , data=data.frame(rugged_std=rugged_seq) )
for ( i in 1:50 ) lines( rugged_seq , mu[i,] , col=col.alpha("black",0.3) )

precis(m8.1)
```

```{r}
## R code 8.6
# Non-African nations
m8.2 <- quap(
    alist(
        log_gdp_std ~ dnorm( mu , sigma ) ,
        mu <- a + b*( rugged_std - 0.215 ) ,
        a ~ dnorm( 1 , 0.1 ) ,
        b ~ dnorm( 0 , 0.25 ) ,
        sigma ~ dexp(1)
    ) ,
    data=d.A0 )

## R code 8.3
set.seed(7)
prior <- extract.prior( m8.2 )

# set up the plot dimensions
plot( NULL , xlim=c(0,1) , ylim=c(0.5,1.5) ,
    xlab="ruggedness" , ylab="log GDP" )
abline( h=min(dd$log_gdp_std) , lty=2 )
abline( h=max(dd$log_gdp_std) , lty=2 )

# draw 50 lines from the prior
rugged_seq <- seq( from=-0.1 , to=1.1 , length.out=30 )
mu <- link( m8.2 , post=prior , data=data.frame(rugged_std=rugged_seq) )
for ( i in 1:50 ) lines( rugged_seq , mu[i,] , col=col.alpha("black",0.3) )

precis(m8.2)
```

### 8.1.2. Adding an indicator variable doesn’t work.

```{r}
## R code 8.7
m8.3 <- quap(
    alist(
        log_gdp_std ~ dnorm( mu , sigma ) ,
        mu <- a + b*( rugged_std - 0.215 ) ,
        a ~ dnorm( 1 , 0.1 ) ,
        b ~ dnorm( 0 , 0.3 ) ,
        sigma ~ dexp( 1 )
    ) ,
    data=dd )

## R code 8.11
precis( m8.3 )
```

```{r}
## R code 8.8
# make variable to index Africa (1) or not (2)
dd$cid <- ifelse( dd$cont_africa==1 , 1 , 2 )

## R code 8.9
m8.4 <- quap(
    alist(
        log_gdp_std ~ dnorm( mu , sigma ) ,
        mu <- a[cid] + b*( rugged_std - 0.215 ) ,
        a[cid] ~ dnorm( 1 , 0.1 ) ,
        b ~ dnorm( 0 , 0.3 ) ,
        sigma ~ dexp( 1 )
    ) ,
    data=dd )

## R code 8.10
compare( m8.3 , m8.4 )

## R code 8.11
precis( m8.4 , depth=2 )
```


```{r}
###
## R code 8.3
set.seed(7)
prior <- extract.prior( m8.4 )

# set up the plot dimensions
plot( NULL , xlim=c(0,1) , ylim=c(0.5,1.5) ,
    xlab="ruggedness" , ylab="log GDP" )
abline( h=min(dd$log_gdp_std) , lty=2 )
abline( h=max(dd$log_gdp_std) , lty=2 )
###

## R code 8.12
rugged.seq <- seq( from=-0.1 , to=1.1 , length.out=30 )

# compute mu over samples, fixing cid=2
mu.NotAfrica <- link( m8.4 ,
    data=data.frame( cid=2 , rugged_std=rugged.seq ) )
for ( i in 1:50 ) lines( rugged_seq , mu.NotAfrica[i,] , col=col.alpha("black",0.3) )

# compute mu over samples, fixing cid=1
mu.Africa <- link( m8.4 ,
    data=data.frame( cid=1 , rugged_std=rugged.seq ) )
for ( i in 1:50 ) lines( rugged_seq , mu.Africa[i,] , col=col.alpha("blue",0.3) )

# summarize to means and intervals
mu.NotAfrica_mu <- apply( mu.NotAfrica , 2 , mean )
mu.NotAfrica_ci <- apply( mu.NotAfrica , 2 , PI , prob=0.97 )
mu.Africa_mu <- apply( mu.Africa , 2 , mean )
mu.Africa_ci <- apply( mu.Africa , 2 , PI , prob=0.97 )

```

### 8.1.3. Adding an interaction does work.

```{r}
## R code 8.13
m8.5 <- quap(
    alist(
        log_gdp_std ~ dnorm( mu , sigma ) ,
        mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
        a[cid] ~ dnorm( 1 , 0.1 ) ,
        b[cid] ~ dnorm( 0 , 0.3 ) ,
        sigma ~ dexp( 1 )
    ) ,
    data=dd )

## R code 8.14
precis( m8.5 , depth=2 )

## R code 8.15
compare( m8.3 , m8.4 , m8.5 )

## R code 8.16
waic_list <- WAIC( m8.5 , pointwise=TRUE )
```

### 8.1.4. Plotting the interaction.

```{r}
## R code 8.17
# plot Africa - cid=1
plot( d.A1$rugged_std , d.A1$log_gdp_std , pch=16 , col=rangi2 ,
    xlab="ruggedness (standardized)" , ylab="log GDP (as proportion of mean)" ,
    xlim=c(0,1) )
mu <- link( m8.5 , data=data.frame( cid=1 , rugged_std=rugged_seq ) )
mu_mean <- apply( mu , 2 , mean )
mu_ci <- apply( mu , 2 , PI , prob=0.97 )
lines( rugged_seq , mu_mean , lwd=2 )
shade( mu_ci , rugged_seq , col=col.alpha(rangi2,0.3) )
mtext("African nations")

# plot non-Africa - cid=2
plot( d.A0$rugged_std , d.A0$log_gdp_std , pch=1 , col="black" ,
    xlab="ruggedness (standardized)" , ylab="log GDP (as proportion of mean)" ,
    xlim=c(0,1) )
mu <- link( m8.5 , data=data.frame( cid=2 , rugged_std=rugged_seq ) )
mu_mean <- apply( mu , 2 , mean )
mu_ci <- apply( mu , 2 , PI , prob=0.97 )
lines( rugged_seq , mu_mean , lwd=2 )
shade( mu_ci , rugged_seq )
mtext("Non-African nations")

```

## 8.5. Practice

### 7E1. For each of the causal relationships below, name a hypothetical third variable that would lead to an interaction effect.

#### (1) Bread dough rises because of yeast.

> 1. tmeperature; 2. the amount or percentage of other ingredients of the dough (like vinegar, flour, sugar)

#### (2) Education leads to higher income.

> 1. family background; 2. work attitudes; 3. location (like which city or country)

#### (3) Gasoline makes a car go.

> 1. what types of car

### 7M3. In parts of North America, ravens depend upon wolves for their food. This is because ravens are carnivorous but cannot usually kill or open carcasses of prey. Wolves however can and do kill and tear open animals, and they tolerate ravens co-feeding at their kills. This species relationship is generally described as a “species interaction.” Can you invent a hypothetical set of data on raven population size in which this relationship would manifest as a statistical interaction? Do you think the biological interaction could be linear? Why or why not?

> Based on the information, I assume that prey population and wolf population interact in predicting raven population. Here is my model:

> raven_pop ~ a1 + beta_prey X prey_pop + beta_wolf X wolf_pop + beta_prey_wolf X prey_pop X wolf_pop

```{r}
N <- 1000 # simulation size
# Simulate data
prey <- rnorm(n = N, mean = 0, sd = 1)

rPW <- 0.6 # correlation between prey and wolf
wolf <- rnorm(n = N, mean = rPW * prey, sd = 1)

bP <- 0.3 # regression coefficient for prey
bW <- 0.1 # regression coefficient for wolf
bPW <- 0.5 # regression coefficient for prey-by-wolf interaction

raven <- rnorm(n = N, mean = bP*prey + bW*wolf + bPW*prey*wolf, sd = 1)

d <- data.frame(raven, prey, wolf)
str(d)
```

```{r}
m7m3 <- quap(
  alist(
    raven ~ dnorm(mu, sigma),
    mu <- a + bP*prey + bW*wolf + bPW*prey*wolf,
    a ~ dnorm(0, 1),
    bW ~ dnorm(0, 1),
    bP ~ dnorm(0, 1),
    bPW ~ dnorm(0, 1),
    sigma ~ dunif(0, 5)
  ),
  data = d)
precis(m7m3)
```



### 7H3. Consider again the data(rugged) data on economic development and terrain ruggedness, examined in this chapter. One of the African countries in that example, Seychelles, is far outside the cloud of other nations, being a rare country with both relatively high GDP and high ruggedness. Seychelles is also unusual, in that it is a group of islands far from the coast of mainland Africa, and its main economic activity is tourism. One might suspect that this one nation is exerting a strong influence on the conclusions. In this problem, I want you to drop Seychelles from the data and re-evaluate the hypothesis that the relationship of African economies with ruggedness is diff erent from that on other continents.

#### (a) Begin by using map to fit just the interaction model:
y i ~ Normal(ui; sigma)
ui = a + bAXAi + bRXRi + bARXAiXRi

#### where y is log GDP per capita in the year 2000 (log of rgdppc_2000); A is cont_africa, the dummy variable for being an African nation; and R is the variable rugged. Choose your own priors. Compare the inference from this model fi t to the data without Seychelles to the same model fit to the full data. Does it still seem like the effect of ruggedness depends upon continent? How much has the expected relationship changed?

```{r}
## R code 8.1
data(rugged)
d <- rugged
# make log version of outcome
d$log_gdp <- log( d$rgdppc_2000 )

# extract countries with GDP data
d <- d[ complete.cases(d$rgdppc_2000) , ]

d2 <- dd[d$country != "Seychelles", ]

# rescale variables
d$log_gdp_std <- d$log_gdp / mean(dd$log_gdp)
d$rugged_std <- d$rugged / max(dd$rugged)
d2$log_gdp_std <- d2$log_gdp / mean(dd$log_gdp)
d2$rugged_std <- d2$rugged / max(dd$rugged)

## R code 8.8
# make variable to index Africa (1) or not (2)
d$cid <- ifelse( d$cont_africa==1 , 1 , 2 )
d2$cid <- ifelse( d2$cont_africa==1 , 1 , 2 )

# split countries into Africa and not-Africa
d.A1 <- d[ d$cont_africa==1 , ] # Africa
d.A0 <- d[ d$cont_africa==0 , ] # not Africa
d2.A1 <- d2[ d2$cont_africa==1 , ] # Africa
d2.A0 <- d2[ d2$cont_africa==0 , ] # not Africa

#str(d)
#str(d2)
```

```{r}
m7H3_with <- quap(
  alist(
    log_gdp_std ~ dnorm( mu , sigma ) ,
    mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
    a[cid] ~ dnorm( 1 , 0.1 ) ,
    b[cid] ~ dnorm( 0 , 0.3 ) ,
    sigma ~ dexp( 1 )
  ),
  data = d)

m7H3_without <- quap(
  alist(
    log_gdp_std ~ dnorm( mu , sigma ) ,
    mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
    a[cid] ~ dnorm( 1 , 0.1 ) ,
    b[cid] ~ dnorm( 0 , 0.3 ) ,
    sigma ~ dexp( 1 )
  ),
  data = d2)

coeftab(m7H3_with, m7H3_without)
```

> a[1] and b[1] decreased. 

#### (b) Now plot the predictions of the interaction model, with and without Seychelles. Does it still seem like the eff ect of ruggedness depends upon continent? How much has the expected relationship changed?

```{r}
par(mfrow = c(2, 2))
##### with
## R code 8.12
rugged.seq <- seq( from=-0.1 , to=1.1 , length.out=30 )
# compute mu over samples, fixing cid=2
mu.NotAfrica <- link( m7H3_with ,
    data=data.frame( cid=2 , rugged_std=rugged.seq ) )
# compute mu over samples, fixing cid=1
mu.Africa <- link( m7H3_with ,
    data=data.frame( cid=1 , rugged_std=rugged.seq ) )
# summarize to means and intervals
mu.NotAfrica_mu <- apply( mu.NotAfrica , 2 , mean )
mu.NotAfrica_ci <- apply( mu.NotAfrica , 2 , PI , prob=0.97 )
mu.Africa_mu <- apply( mu.Africa , 2 , mean )
mu.Africa_ci <- apply( mu.Africa , 2 , PI , prob=0.97 )

## R code 8.17
# plot Africa - cid=1
plot( d.A1$rugged_std , d.A1$log_gdp_std , pch=16 , col=rangi2 ,
    xlab="ruggedness (standardized)" , ylab="log GDP (as proportion of mean)" ,
    xlim=c(0,1) , ylim=c(0.6,1.4))
mu <- link( m7H3_with , data=data.frame( cid=1 , rugged_std=rugged_seq ) )
mu_mean <- apply( mu , 2 , mean )
mu_ci <- apply( mu , 2 , PI , prob=0.97 )
lines( rugged_seq , mu_mean , lwd=2 )
shade( mu_ci , rugged_seq , col=col.alpha(rangi2,0.3) )
mtext("African nations_with")

# plot non-Africa - cid=2
plot( d.A0$rugged_std , d.A0$log_gdp_std , pch=1 , col="black" ,
    xlab="ruggedness (standardized)" , ylab="log GDP (as proportion of mean)" ,
    xlim=c(0,1) , ylim=c(0.6,1.4))
mu <- link( m7H3_with , data=data.frame( cid=2 , rugged_std=rugged_seq ) )
mu_mean <- apply( mu , 2 , mean )
mu_ci <- apply( mu , 2 , PI , prob=0.97 )
lines( rugged_seq , mu_mean , lwd=2 )
shade( mu_ci , rugged_seq )
mtext("Non-African nations_with")

##### without

## R code 8.12
rugged.seq <- seq( from=-0.1 , to=1.1 , length.out=30 )
# compute mu over samples, fixing cid=2
mu.NotAfrica <- link( m7H3_without ,
    data=data.frame( cid=2 , rugged_std=rugged.seq ) )
# compute mu over samples, fixing cid=1
mu.Africa <- link( m7H3_without ,
    data=data.frame( cid=1 , rugged_std=rugged.seq ) )
# summarize to means and intervals
mu.NotAfrica_mu <- apply( mu.NotAfrica , 2 , mean )
mu.NotAfrica_ci <- apply( mu.NotAfrica , 2 , PI , prob=0.97 )
mu.Africa_mu <- apply( mu.Africa , 2 , mean )
mu.Africa_ci <- apply( mu.Africa , 2 , PI , prob=0.97 )

## R code 8.17
# plot Africa - cid=1
plot( d2.A1$rugged_std , d2.A1$log_gdp_std , pch=16 , col=rangi2 ,
    xlab="ruggedness (standardized)" , ylab="log GDP (as proportion of mean)" ,
    xlim=c(0,1) , ylim=c(0.6,1.4))
mu <- link( m7H3_without , data=data.frame( cid=1 , rugged_std=rugged_seq ) )
mu_mean <- apply( mu , 2 , mean )
mu_ci <- apply( mu , 2 , PI , prob=0.97 )
lines( rugged_seq , mu_mean , lwd=2 )
shade( mu_ci , rugged_seq , col=col.alpha(rangi2,0.3) )
mtext("African nations_without")

# plot non-Africa - cid=2
plot( d2.A0$rugged_std , d2.A0$log_gdp_std , pch=1 , col="black" ,
    xlab="ruggedness (standardized)" , ylab="log GDP (as proportion of mean)" ,
    xlim=c(0,1) , ylim=c(0.6,1.4))
mu <- link( m7H3_without , data=data.frame( cid=2 , rugged_std=rugged_seq ) )
mu_mean <- apply( mu , 2 , mean )
mu_ci <- apply( mu , 2 , PI , prob=0.97 )
lines( rugged_seq , mu_mean , lwd=2 )
shade( mu_ci , rugged_seq )
mtext("Non-African nations_without")
```

> Without Seychelles, the slope of African nations_without is a bit shallower. The uncertainty of African nations_without at higher ruggedness is larger.

#### Method 2 (fit ui = a + bAXAi + bRXRi + bARXAiXRi)

```{r}
d <- rugged[complete.cases(rugged$rgdppc_2000), ]
d$log_gdp <- log(d$rgdppc_2000)
d2 <- d[d$country != "Seychelles", ]
m_with <- quap(
  alist(
    log_gdp ~ dnorm(mu, sigma),
    mu <- a + bA*cont_africa + bR*rugged + bAR*cont_africa*rugged,
    a ~ dnorm(8, 100),
    bA ~ dnorm(0, 1),
    bR ~ dnorm(0, 1),
    bAR ~ dnorm(0, 1),
    sigma ~ dunif(0, 10)
  ),
  data = d
)
m_without <- quap(
  alist(
    log_gdp ~ dnorm(mu, sigma),
    mu <- a + bA*cont_africa + bR*rugged + bAR*cont_africa*rugged,
    a ~ dnorm(8, 100),
    bA ~ dnorm(0, 1),
    bR ~ dnorm(0, 1),
    bAR ~ dnorm(0, 1),
    sigma ~ dunif(0, 10)
  ),
  data = d2
)
coeftab(m_with, m_without)
```

```{r}
par(mfrow = c(2, 2))
# Predictions for model with Seychelles
rugged.seq <- seq(from = -1, to = 8, by = 0.25)
mu.Africa <- link(m_with, data = data.frame(cont_africa = 1, rugged = rugged.seq), refresh = 0)
mu.Africa.mean <- apply(mu.Africa, 2, mean)
mu.Africa.PI <- apply(mu.Africa, 2, PI, prob = 0.97)
mu.NotAfrica <- link(m_with, data = data.frame(cont_africa = 0, rugged = rugged.seq), refresh = 0)
mu.NotAfrica.mean <- apply(mu.NotAfrica, 2, mean)
mu.NotAfrica.PI <- apply(mu.NotAfrica, 2, PI, prob = 0.97)
d.A1 <- d[d$cont_africa == 1, ]
plot(log_gdp ~ rugged, data = d.A1, col = rangi2)
mtext("African nations, with Seychelles", 3)
lines(rugged.seq, mu.Africa.mean, col = rangi2)
shade(mu.Africa.PI, rugged.seq, col = col.alpha(rangi2, 0.3))

d.A0 <- d[d$cont_africa == 0, ]
plot(log_gdp ~ rugged, data = d.A0)
mtext("Non-African nations, with Seychelles", 3)
lines(rugged.seq, mu.NotAfrica.mean)
shade(mu.NotAfrica.PI, rugged.seq)


# Predictions for model without Seychelles
rugged.seq <- seq(from = -1, to = 8, by = 0.25)
mu.Africa <- link(m_without, data = data.frame(cont_africa = 1, rugged = rugged.seq), refresh = 0)
mu.Africa.mean <- apply(mu.Africa, 2, mean)
mu.Africa.PI <- apply(mu.Africa, 2, PI, prob = 0.97)
mu.NotAfrica <- link(m_without, data = data.frame(cont_africa = 0, rugged = rugged.seq), refresh = 0)
mu.NotAfrica.mean <- apply(mu.NotAfrica, 2, mean)
mu.NotAfrica.PI <- apply(mu.NotAfrica, 2, PI, prob = 0.97)
d.A1 <- d2[d2$cont_africa == 1, ]
plot(log_gdp ~ rugged, data = d.A1, col = rangi2)
mtext("African nations, without Seychelles", 3)
lines(rugged.seq, mu.Africa.mean, col = rangi2)
shade(mu.Africa.PI, rugged.seq, col = col.alpha(rangi2, 0.3))

d.A0 <- d2[d2$cont_africa == 0, ]
plot(log_gdp ~ rugged, data = d.A0)
mtext("Non-African nations, without Seychelles", 3) 
lines(rugged.seq, mu.NotAfrica.mean)
shade(mu.NotAfrica.PI, rugged.seq)
```



#### (c) Finally, conduct a model comparison analysis, using WAIC. Fit three models to the data without Seychelles:

Model 1 : y i ~ Normal(ui; sigma)
ui = a + bRXRi


Model 2 : y i ~ Normal(ui; sigma)
ui = a + bAXAi + bRXRi


Model 3 : y i ~ Normal(ui; sigma)
ui = a + bAXAi + bRXRi + bARXAiXRi

#### Use whatever priors you think are sensible. Plot the model-averaged predictions of this model set. Do your inferences diff er from those in (b)? Why or why not?

```{r}
m_with1 <- quap(
  alist(
    log_gdp ~ dnorm(mu, sigma),
    mu <- a + bR*rugged,
    a ~ dnorm(8, 100),
    bR ~ dnorm(0, 1),
    sigma ~ dunif(0, 10)
  ),
  data = d
)
m_with2 <- quap(
  alist(
    log_gdp ~ dnorm(mu, sigma),
    mu <- a + bA*cont_africa + bR*rugged,
    a ~ dnorm(8, 100),
    bA ~ dnorm(0, 1),
    bR ~ dnorm(0, 1),
    sigma ~ dunif(0, 10)
  ),
  data = d
)
m_with3 <- quap(
  alist(
    log_gdp ~ dnorm(mu, sigma),
    mu <- a + bA*cont_africa + bR*rugged + bAR*cont_africa*rugged,
    a ~ dnorm(8, 100),
    bA ~ dnorm(0, 1),
    bR ~ dnorm(0, 1),
    bAR ~ dnorm(0, 1),
    sigma ~ dunif(0, 10)
  ),
  data = d
)
compare(m_with1, m_with2, m_with3)
```

```{r}
m_without1 <- map(
  alist(
    log_gdp ~ dnorm(mu, sigma),
    mu <- a + bR*rugged,
    a ~ dnorm(8, 100),
    bR ~ dnorm(0, 1),
    sigma ~ dunif(0, 10)
  ),
  data = d2
)
m_without2 <- map(
  alist(
    log_gdp ~ dnorm(mu, sigma),
    mu <- a + bA*cont_africa + bR*rugged,
    a ~ dnorm(8, 100),
    bA ~ dnorm(0, 1),
    bR ~ dnorm(0, 1),
    sigma ~ dunif(0, 10)
  ),
  data = d2
)
m_without3 <- map(
  alist(
    log_gdp ~ dnorm(mu, sigma),
    mu <- a + bA*cont_africa + bR*rugged + bAR*cont_africa*rugged,
    a ~ dnorm(8, 100),
    bA ~ dnorm(0, 1),
    bR ~ dnorm(0, 1),
    bAR ~ dnorm(0, 1),
    sigma ~ dunif(0, 10)
  ),
  data = d2
)
compare(m_without1, m_without2, m_without3)
```

> Without Seychelles, m_without3 has less weight than m_with3. This indicates that Seychelles affects the interaction term.


#### 1. Use the tomato.csv (attached) data set and evaluate whether hypocotyl length ("hyp") is affected by shade ("trt"), species ("species") and their interaction.

```{r}
tomato <- read.csv("Tomato.csv")
str(tomato)
head(tomato)
```

```{r}
tomato2 <- subset(tomato, select = c(hyp, trt, species))
str(tomato2)
head(tomato2)
```

```{r}

```

