---
title: "chapter 13"
author: "Min-Yao"
date: "2020/1/7"
output: 
  html_document: 
    keep_md: yes
---

# 13 Models With Memory
## 13.1. Example: Multilevel tadpoles
```{r}
## R code 13.1
library(rethinking)
data(reedfrogs)
d <- reedfrogs
str(d)

## R code 13.2
# make the tank cluster variable
d$tank <- 1:nrow(d)

dat <- list(
    S = d$surv,
    N = d$density,
    tank = d$tank )

# approximate posterior
m13.1 <- ulam(
    alist(
        S ~ dbinom( N , p ) ,
        logit(p) <- a[tank] ,
        a[tank] ~ dnorm( 0 , 1.5 )
    ), data=dat , chains=4 , log_lik=TRUE )

## R code 13.3
m13.2 <- ulam(
    alist(
        S ~ dbinom( N , p ) ,
        logit(p) <- a[tank] ,
        a[tank] ~ dnorm( a_bar , sigma ) ,
        a_bar ~ dnorm( 0 , 1.5 ) ,
        sigma ~ dexp( 1 )
    ), data=dat , chains=4 , log_lik=TRUE )

## R code 13.4
compare( m13.1 , m13.2 )

## R code 13.5
# extract Stan samples
post <- extract.samples(m13.2)

# compute median intercept for each tank
# also transform to probability with logistic
d$propsurv.est <- logistic( apply( post$a , 2 , mean ) )

# display raw proportions surviving in each tank
plot( d$propsurv , ylim=c(0,1) , pch=16 , xaxt="n" ,
    xlab="tank" , ylab="proportion survival" , col=rangi2 )
axis( 1 , at=c(1,16,32,48) , labels=c(1,16,32,48) )

# overlay posterior means
points( d$propsurv.est )

# mark posterior mean probability across tanks
abline( h=mean(inv_logit(post$a_bar)) , lty=2 )

# draw vertical dividers between tank densities
abline( v=16.5 , lwd=0.5 )
abline( v=32.5 , lwd=0.5 )
text( 8 , 0 , "small tanks" )
text( 16+8 , 0 , "medium tanks" )
text( 32+8 , 0 , "large tanks" )

## R code 13.6
# show first 100 populations in the posterior
plot( NULL , xlim=c(-3,4) , ylim=c(0,0.35) ,
    xlab="log-odds survive" , ylab="Density" )
for ( i in 1:100 )
    curve( dnorm(x,post$a_bar[i],post$sigma[i]) , add=TRUE ,
    col=col.alpha("black",0.2) )

# sample 8000 imaginary tanks from the posterior distribution
sim_tanks <- rnorm( 8000 , post$a_bar , post$sigma )

# transform to probability and visualize
dens( inv_logit(sim_tanks) , lwd=2 , adj=0.1 )
```

## 13.2. Varying eff ects and the underfi tting/overfi tting trade-off
### 13.2.1. The model.
### 13.2.2. Assign values to the parameters.
```{r}
## R code 13.7
a_bar <- 1.5
sigma <- 1.5
nponds <- 60
Ni <- as.integer( rep( c(5,10,25,35) , each=15 ) )

## R code 13.8
set.seed(5005)
a_pond <- rnorm( nponds , mean=a_bar , sd=sigma )

## R code 13.9
dsim <- data.frame( pond=1:nponds , Ni=Ni , true_a=a_pond )

## R code 13.10
class(1:3)
class(c(1,2,3))
```

### 13.2.3. Simulate survivors.

```{r}
## R code 13.11
dsim$Si <- rbinom( nponds , prob=logistic(dsim$true_a) , size=dsim$Ni )
```

### 13.2.4. Compute the no-pooling estimates.
```{r}
## R code 13.12
dsim$p_nopool <- dsim$Si / dsim$Ni
```

### 13.2.5. Compute the partial-pooling estimates.
```{r}
## R code 13.13
dat <- list( Si=dsim$Si , Ni=dsim$Ni , pond=dsim$pond )
m13.3 <- ulam(
    alist(
        Si ~ dbinom( Ni , p ),
        logit(p) <- a_pond[pond],
        a_pond[pond] ~ dnorm( a_bar , sigma ),
        a_bar ~ dnorm( 0 , 1.5 ),
        sigma ~ dexp( 1 )
    ), data=dat , chains=4 )

## R code 13.14
precis( m13.3 , depth=2 )

## R code 13.15
post <- extract.samples( m13.3 )
dsim$p_partpool <- apply( inv_logit(post$a_pond) , 2 , mean )

## R code 13.16
dsim$p_true <- inv_logit( dsim$true_a )

## R code 13.17
nopool_error <- abs( dsim$p_nopool - dsim$p_true )
partpool_error <- abs( dsim$p_partpool - dsim$p_true )

## R code 13.18
plot( 1:60 , nopool_error , xlab="pond" , ylab="absolute error" ,
    col=rangi2 , pch=16 )
points( 1:60 , partpool_error )

## R code 13.19
nopool_avg <- aggregate(nopool_error,list(dsim$Ni),mean)
partpool_avg <- aggregate(partpool_error,list(dsim$Ni),mean)

## R code 13.20
a <- 1.5
sigma <- 1.5
nponds <- 60
Ni <- as.integer( rep( c(5,10,25,35) , each=15 ) )
a_pond <- rnorm( nponds , mean=a , sd=sigma )
dsim <- data.frame( pond=1:nponds , Ni=Ni , true_a=a_pond )
dsim$Si <- rbinom( nponds,prob=inv_logit( dsim$true_a ),size=dsim$Ni )
dsim$p_nopool <- dsim$Si / dsim$Ni
newdat <- list(Si=dsim$Si,Ni=dsim$Ni,pond=1:nponds)
m13.3new <- stan( fit=m13.3@stanfit , data=newdat , chains=4 )

post <- extract.samples( m13.3new )
dsim$p_partpool <- apply( inv_logit(post$a_pond) , 2 , mean )
dsim$p_true <- inv_logit( dsim$true_a )
nopool_error <- abs( dsim$p_nopool - dsim$p_true )
partpool_error <- abs( dsim$p_partpool - dsim$p_true )
plot( 1:60 , nopool_error , xlab="pond" , ylab="absolute error" , col=rangi2 , pch=16 )
points( 1:60 , partpool_error )
```

## 13.6. Summary

## 13.7. Practice

###Easy.

### 12E1. Which of the following priors will produce more shrinkage in the estimates? (a) αtank ∼ Normal(0, 1); (b) αtank ∼ Normal(0, 2).

### 12E2. Make the following model into a multilevel model.
y i ∼ Binomial(1, p i)
logit(p i) = α
group[i] + βx i
αgroup ∼ Normal(0, 10)
β ∼ Normal(0, 1)

### 12E3. Make the following model into a multilevel model.
y i ∼ Normal(µi, σ)
µi = α
group[i] + βx i
αgroup ∼ Normal(0, 10)
β ∼ Normal(0, 1)
σ ∼ HalfCauchy(0, 2)

### Medium.

### 12M1. Revisit the Reed frog survival data, data(reedfrogs), and add the predation and size treatment variables to the varying intercepts model. Consider models with either main eff ect alone, both main eff ects, as well as a model including both and their interaction. Instead of focusing on inferences about these two predictor variables, focus on the inferred variation across tanks. Explain why it changes as it does across models.

### 12M2. Compare the models you fi t just above, using WAIC. Can you reconcile the differences in WAIC with the posterior distributions of the models?

### 12H1. In 1980, a typical Bengali woman could have 5 or more children in her lifetime. By the year 200, a typical Bengali woman had only 2 or 3. You’re going to look at a historical set of data, when contraception was widely available but many families chose not to use it. Th ese data reside in data(bangladesh) and come from the 1988 Bangladesh Fertility Survey. Each row is one of 1934 women. Th ere are six variables, but you can focus on three of them for this practice problem:

(1) district: ID number of administrative district each woman resided in
(2) use.contraception: An indicator (0/1) of whether the woman was using contraception
(3) urban: An indicator (0/1) of whether the woman lived in a city, as opposed to living in a
rural area

The first thing to do is ensure that the cluster variable, district, is a contiguous set of integers. Recall that these values will be index values inside the model. If there are gaps, you’ll have parameters for which there is no data to inform them. Worse, the model probably won’t run. Look at the unique values of the district variable:
```{r}
## R code 13.40
sort(unique(d$district))
```

District 54 is absent. So district isn’t yet a good index variable, because it’s not contiguous. Th is is
easy to fi x. Just make a new variable that is contiguous. Th is is enough to do it:
```{r}
## R code 13.41
d$district_id <- as.integer(as.factor(d$district))
sort(unique(d$district_id))
```

Now there are 60 values, contiguous integers 1 to 60.
Now, focus on predicting use.contraception, clustered by district_id. Do not include
urban just yet. Fit both (1) a traditional fi xed-eff ects model that uses dummy variables for district and
(2) a multilevel model with varying intercepts for district. Plot the predicted proportions of women
in each district using contraception, for both the fi xed-eff ects model and the varying-eff ects model.
Th at is, make a plot in which district ID is on the horizontal axis and expected proportion using con-
traception is on the vertical. Make one plot for each model, or layer them on the same plot, as you
prefer. How do the models disagree? Can you explain the pattern of disagreement? In particular, can
you explain the most extreme cases of disagreement, both why they happen where they do and why
the models reach diff erent inferences?