---
title: "chapter 2"
author: "Min-Yao"
date: "2019¦~4¤ë11¤é"
output: 
  html_document: 
    keep_md: yes
---

```{r}
library("rstan") # observe startup messages
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
Sys.setenv(LOCAL_CPPFLAGS = '-march=native')
```

# Chapter 0

## R code 0.1
```{r}
print( "All models are wrong, but some are useful." )
```

## R code 0.2
```{r}
x <- 1:2
x
x <- x*10
x
x <- log(x)
x
x <- sum(x)
x
x <- exp(x)
x
```

## R code 0.3
```{r}
( log( 0.01^200 ) )
( 200 * log(0.01) )
```

## R code 0.4
#### Load the data:
#### car braking distances in feet paired with speeds in km/h
#### see ?cars for details
```{r}
data(cars)
```


#### fit a linear regression of distance on speed
```{r}
m <- lm( dist ~ speed , data=cars )
```

#### estimated coefficients from the model
```{r}
coef(m)
```


#### plot residuals against speed
```{r}
plot( resid(m) ~ speed , data=cars )
```


## R code 0.5
```{r}
#install.packages(c("coda","mvtnorm","devtools"))
library(devtools)
#devtools::install_github("rmcelreath/rethinking",ref="Experimental")
```

# Chapter 2

## R code 2.1

```{r}
ways <- c( 0 , 3 , 8 , 9 , 0 )
ways/sum(ways)
```

## R code 2.2
```{r}
dbinom( 6 , size=9 , prob=0.5 )
```


## R code 2.3
#### define grid
```{r}
p_grid <- seq( from=0 , to=1 , length.out=20 )
```


#### define prior
```{r}
prior <- rep( 1 , 20 )
```


#### compute likelihood at each value in grid
```{r}
likelihood <- dbinom( 6 , size=9 , prob=p_grid )
```

#### compute product of likelihood and prior
```{r}
unstd.posterior <- likelihood * prior
```

#### standardize the posterior, so it sums to 1
```{r}
posterior <- unstd.posterior / sum(unstd.posterior)
```

## R code 2.4
```{r}
plot( p_grid , posterior , type="b" ,
    xlab="probability of water" , ylab="posterior probability" )
mtext( "20 points" )
```

## R code 2.5
```{r}
prior <- ifelse( p_grid < 0.5 , 0 , 1 )
prior <- exp( -5*abs( p_grid - 0.5 ) )
```

## R code 2.6
```{r}
library(rethinking)
library(parallel)
globe.qa <- quap(
    alist(
        W ~ dbinom( W+L ,p) ,  # binomial likelihood
        p ~ dunif(0,1)     # uniform prior
    ) ,
    data=list(W=6,L=3) )
```

# display summary of quadratic approximation
```{r}
precis( globe.qa )
```


## R code 2.7

```{r}
# analytical calculation
W <- 6
L <- 3
curve( dbeta( x , W+1 , L+1 ) , from=0 , to=1 )

# quadratic approximation
curve( dnorm( x , 0.67 , 0.16 ) , lty=2 , add=TRUE )
```


## R code 2.8
```{r}
n_samples <- 1000
p <- rep( NA , n_samples )
p[1] <- 0.5
W <- 6
L <- 3
for ( i in 2:n_samples ) {
    p_new <- rnorm( 1 , p[i-1] , 0.1 )
    if ( p_new < 0 ) p_new <- abs( p_new )
    if ( p_new > 1 ) p_new <- 2 - p_new
    q0 <- dbinom( W , W+L , p[i-1] )
    q1 <- dbinom( W , W+L , p_new )
    p[i] <- ifelse( runif(1) < q1/q0 , p_new , p[i-1] )
}
```

## R code 2.9
```{r}
dens( p , xlim=c(0,1) )
curve( dbeta( x , W+1 , L+1 ) , lty=2 , add=TRUE )
```

### 2E1 Which of the expressions below correspond to the statement: the probability of rain on Monday?

(1) Pr(rain)


### 2E2

(3) Th e probability that it is Monday, given that it is raining.


### 2E3

(1) Pr(Mondayjrain)
(4) Pr(rainjMonday) Pr(Monday)= Pr(rain)


### 2M1-(1)
```{r}
data1 <- c('W','W','W')
# total number of tosses
n = length(data1) 
# observed number of water
w = sum(data1=='W')  

# define grid
p_grid <- seq( from=0 , to=1 , length.out=100)
length_of_grid = length(p_grid)
# prior probability of p
prior <- rep( 1 , length_of_grid ) 

# compute likelihood at each value in grid
likelihood <- dbinom( w , size=n , prob=p_grid )

# compute product of likelihood and prior
unstd.posterior <- likelihood * prior

# standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)

plot_posterior <- function(x, y) {
  plot(x = x, y = y, type="b", xlab = "Probability of Water", ylab = "Posterior Probability")
  title <- paste( length(x), "Points")
  mtext(title)
}
plot_posterior(x = p_grid, y = posterior)

```
### 2M1-(2)
```{r}
data1 <- c('W','W','W','L')
# total number of tosses
n = length(data1) 
# observed number of water
w = sum(data1=='W')  

# define grid
p_grid <- seq( from=0 , to=1 , length.out=100)
length_of_grid = length(p_grid)
# prior probability of p
prior <- rep( 1 , length_of_grid ) 

# compute likelihood at each value in grid
likelihood <- dbinom( w , size=n , prob=p_grid )

# compute product of likelihood and prior
unstd.posterior <- likelihood * prior

# standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)

plot_posterior(x = p_grid, y = posterior)

```

### 2M1-(3)
```{r}
data1 <- c('L', 'W', 'W', 'L', 'W', 'W', 'W')
# total number of tosses
n = length(data1)
# observed number of water
w = sum(data1=='W') 

# define grid
p_grid <- seq( from=0 , to=1 , length.out=100)
length_of_grid = length(p_grid)
# prior probability of p
prior <- rep( 1 , length_of_grid ) 

# compute likelihood at each value in grid
likelihood <- dbinom( w , size=n , prob=p_grid )

# compute product of likelihood and prior
unstd.posterior <- likelihood * prior

# standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)

plot_posterior(x = p_grid, y = posterior)

```

### 2M2-(1)
```{r}
data1 <- c('W','W','W')
# total number of tosses
n = length(data1) 
# observed number of water
w = sum(data1=='W')  

# define grid
p_grid <- seq( from=0 , to=1 , length.out=100)
length_of_grid = length(p_grid)

# prior probability of p
prior <- ifelse( p_grid < 0.5 , 0 , 1 ) 

# compute likelihood at each value in grid
likelihood <- dbinom( w , size=n , prob=p_grid )

# compute product of likelihood and prior
unstd.posterior <- likelihood * prior

# standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)

plot_posterior <- function(x, y) {
  plot(x = x, y = y, type="b", xlab = "Probability of Water", ylab = "Posterior Probability")
  title <- paste( length(x), "Points")
  mtext(title)
}
plot_posterior(x = p_grid, y = posterior)

```

### 2M2-(2)
```{r}
data1 <- c('W','W','W','L')
# total number of tosses
n = length(data1) 
# observed number of water
w = sum(data1=='W')  

# define grid
p_grid <- seq( from=0 , to=1 , length.out=100)
length_of_grid = length(p_grid)

# prior probability of p
prior <- ifelse( p_grid < 0.5 , 0 , 1 ) 

# compute likelihood at each value in grid
likelihood <- dbinom( w , size=n , prob=p_grid )

# compute product of likelihood and prior
unstd.posterior <- likelihood * prior

# standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)

plot_posterior(x = p_grid, y = posterior)

```

### 2M2-(3)
```{r}
data1 <- c('L', 'W', 'W', 'L', 'W', 'W', 'W')
# total number of tosses
n = length(data1)
# observed number of water
w = sum(data1=='W') 

# define grid
p_grid <- seq( from=0 , to=1 , length.out=100)
length_of_grid = length(p_grid)

# prior probability of p
prior <- ifelse( p_grid < 0.5 , 0 , 1 )  

# compute likelihood at each value in grid
likelihood <- dbinom( w , size=n , prob=p_grid )

# compute product of likelihood and prior
unstd.posterior <- likelihood * prior

# standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)

plot_posterior(x = p_grid, y = posterior)

```

### 2M3

```{r}
# p(Earth) and p(Mars)
p_Earth = 0.5  
p_Mars = 0.5 

# p(land|Earth) and p(land|Mars)
p_land_E = 0.3  
p_land_M = 1.0 

# p(land)
p_land = p_Earth*p_land_E + p_Mars*p_land_M
p_land

# p(Earth|land) = (p(land|Earth) * p(Earth)) / p(land)
p_Earth_land = (p_land_E * p_Earth)/p_land 
p_Earth_land


```

### 2M4
```{r}
# Use the counting method
BB.likelihood <- 2
BW.likelihood <- 1
WW.likelihood <- 0

likelihood <- c(BB.likelihood, BW.likelihood, WW.likelihood)

# prior probability
prior <- rep(1, length = length(likelihood))

# compute product of likelihood and prior
unstd.posterior <- likelihood * prior

# standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)
posterior

posterior[1] == 2/3
```



### 2M5
```{r}
# Use the counting method
BB.likelihood <- 2
BW.likelihood <- 1
WW.likelihood <- 0

likelihood2 <- c(BB.likelihood, BW.likelihood, WW.likelihood, BB.likelihood)

# prior probability
prior <- rep(1, length = length(likelihood2))

# compute product of likelihood and prior
unstd.posterior <- likelihood2 * prior

# standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)
posterior

# we may draw card 1 or 4
result = posterior[1]+posterior[4]
result
```

