---
title: "brms_intro_HW"
author: "Min-Yao"
date: "2020/8/16"
output: 
  html_document: 
    keep_md: yes
---

# Assignment


```{r}
library(rethinking)
library(brms)
library(tidyverse)
```

Revisit the following homework problems and try to fit the with brms.  Make your first attempt without looking at the rethinking to brms translation, but if you get stuck definitely look!  Compare the coefficients or predictions that you obtain with brms and those with quap or ulam.

* 4H1, 4H2 (you probably need the function `posterior_predict()`)

### 4H1, The weights listed below were recorded in the !Kung census, but heights were not recorded for these individuals. Provide predicted heights and 89% intervals (either HPDI or PI) for each of these individuals. That is, fill in the table below, using model-based predictions.

Individual weight expected height 89% interval
1 46.95
2 43.72
3 64.78
4 32.59
5 54.63

```{r}
weights <- c(46.95,43.72,64.78,32.59,54.63)
```


```{r}
## change from R code 4.42
# load data again, since it's a long way back
library(rethinking)
data(Howell1)
d <- Howell1
str(d)
```

```{r}
d2 <- d
str(d2)
m4H1brm <- brm(height ~ weight, data=d2,
               prior=c(
                 set_prior("normal(178, 20)", class = "Intercept"),
                 set_prior("lognormal(0,1)", class = "b",lb=0),
                 set_prior("exponential(1)", class = "sigma"))
               )
```
```{r}
m4H1brm
```
```{r}
plot(m4H1brm)
```

```{r}
## predicted responses
p_m4H1brm <- predict(m4H1brm, newdata = data.frame(weight=weights))
head(p_m4H1brm)
```
```{r}
pp_m4H1brm <- posterior_predict(m4H1brm, newdata = data.frame(weight=weights))
head(pp_m4H1brm)
```

### 4H2 Select out all the rows in the Howell1 data with ages below 18 years of age. If you do it right, you should end up with a new data frame with 192 rows in it.

```{r}
## change from R code 4.42
# load data again, since it's a long way back
library(rethinking)
data(Howell1)
d <- Howell1
d4H2 <- d[ d$age < 18 , ]
str(d4H2)
```
#### (a) Fit a linear regression to these data, using quap. Present and interpret the estimates. For every 10 units of increase in weight, how much taller does the model predict a child gets?

```{r}
m4H2brm <- brm(height ~ weight, data=d4H2,
               prior=c(
                 set_prior("normal(178, 20)", class = "Intercept"),
                 set_prior("lognormal(0,1)", class = "b",lb=0),
                 set_prior("exponential(1)", class = "sigma"))
               )
```
```{r}
m4H2brm
plot(m4H2brm)
```
> For every 10 units of increase in weight, the model predict a child gets 27.2 cm taller.

#### (b) Plot the raw data, with height on the vertical axis and weight on the horizontal axis. Super-impose the MAP regression line and 89% HPDI for the mean. Also superimpose the 89% HPDI for predicted heights.

```{r}
## change from R code 4.54
# define sequence of weights to compute predictions for
# these values will be on the horizontal axis
weight.seq <- seq( from= min(d4H2$weight) , to= max(d4H2$weight) , by=1 )

## predicted responses
p_m4H2brm <- predict(m4H2brm, newdata = data.frame(weight=weight.seq))
head(p_m4H2brm)
df_p_m4H2brm <- data.frame(p_m4H2brm)
str(df_p_m4H2brm)
```

```{r}
## change from R code 4.61
# plot raw data
plot( height ~ weight , d4H2 , col=col.alpha(rangi2,1) )
# draw MAP line
lines( weight.seq , df_p_m4H2brm$Estimate )
# draw HPDI region for line
lines( weight.seq , df_p_m4H2brm$Q2.5 )
lines( weight.seq , df_p_m4H2brm$Q97.5 )

```


#### (c) What aspects of the model fit concern you? Describe the kinds of assumptions you would change, if any, to improve the model. You don’t have to write any new code. Just explain what the model appears to be doing a bad job of, and what you hypothesize would be a better model.

> I think a parabolic (second order) polynomial would be a better model.



* From chapter 8 I assigned a tomato problem from my data "Use the tomato.csv data set and evaluate whether hypocotyl length ("hyp") is affected by shade ("trt"), species ("species") and their interaction."

#### 1. Use the tomato.csv (attached) data set and evaluate whether hypocotyl length ("hyp") is affected by shade ("trt"), species ("species") and their interaction.

```{r}
tomato <- read.csv("Tomato.csv")
str(tomato)
head(tomato)
```

```{r}
tomato2 <- subset(tomato, select = c(hyp, trt, species))
str(tomato2)
head(tomato2)
```

```{r}
tomato2$SID <- as.numeric(as.factor(tomato2$species))
tomato2$TID <- as.numeric(as.factor(tomato2$trt))-1
str(tomato2)
head(tomato2)
summary(tomato2)
```
#### no interaction model
##### rethinking
```{r}
tomato_m1 <- quap(
  alist(
    hyp ~ dnorm( mu , sigma ) ,
    mu <- a[SID] + b*TID ,
    a[SID] ~ dnorm( 30 , 5 ) ,
    b ~ dnorm( 0 , 1 ) ,
    sigma ~ dexp( 1 )) ,
    data=tomato2 )
precis(tomato_m1, depth=2)
```

```{r}
### interaction model
tomato_m2 <- quap(
  alist(
    hyp ~ dnorm( mu , sigma ) ,
    mu <- a[SID] + b*TID + bST[SID]*TID,
    a[SID] ~ dnorm( 30 , 5 ) ,
    b ~ dnorm( 0 , 1 ) ,
    bST[SID] ~ dnorm( 0 , 1 ),
    sigma ~ dexp( 1 )) ,
    data=tomato2 )

precis(tomato_m2, depth=2)
```



##### brms


```{r}
get_prior(hyp ~ 1 + SID + TID, data=tomato2)
get_prior(hyp ~ 0 + SID + TID, data=tomato2)
```

```{r}
tomato_m1_brms <-
  brm(data = tomato2, 
      family = gaussian,
      hyp ~ 1 + TID,
      prior = c(prior(normal(30 , 5), class = Intercept),
                prior(normal(0 , 1), class = b),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 8)
```
```{r}
tomato_m1_brms
```
```{r}
# tomato_m1_brms_2 <-
#   brm(data = tomato2, 
#       family = gaussian,
#       hyp ~ 0 + SID + TID,
#       prior = c(prior(normal(30 , 5), class = b, coef = SID1),
#                 prior(normal(30 , 5), class = b, coef = SID2),
#                 prior(normal(30 , 5), class = b, coef = SID3),
#                 prior(normal(30 , 5), class = b, coef = SID4),
#                 prior(normal(30 , 5), class = b, coef = SID5),
#                 prior(normal(0 , 1), class = b),
#                 prior(exponential(1), class = sigma)),
#       iter = 2000, warmup = 1000, chains = 4, cores = 4,
#       seed = 8)

tomato_m1_brms_2 <-
  brm(data = tomato2, 
      family = gaussian,
      hyp ~ 0 + SID + TID,
      prior = c(prior(normal(30 , 5), class = b , coef = SID),
                prior(normal(0 , 1), class = b, coef = TID),
                prior(exponential(1), class = sigma)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 8)
```
```{r}
tomato_m1_brms_2
```

* From chapter 9: 8M1 (remember that the problem numbers were offset it is actually called 9M1 in the Nov 24 PDF)

#### 8M1. Re-estimate the terrain ruggedness model from the chapter, but now using a uniform prior and an exponential prior for the standard deviation, sigma. The uniform prior should be dunif(0,10) and the exponential should be dexp(1). Do the different priors have any detectible influence on the posterior distribution?

```{r}
## R code 9.9
library(rethinking)
data(rugged)
d <- rugged
```


```{r}
d <- 
  d %>%
  mutate(log_gdp = log(rgdppc_2000))

dd <-
  d %>%
  drop_na(rgdppc_2000) %>% 
  mutate(log_gdp_std = log_gdp / mean(log_gdp),
         rugged_std  = rugged / max(rugged),
         cid         = ifelse(cont_africa == 1, "1", "2")) %>% 
  mutate(rugged_std_c = rugged_std - mean(rugged_std))
```

#### exponential prior

```{r}
b9.1 <- 
  brm(data = dd, 
      family = gaussian,
      bf(log_gdp_std ~ 0 + a + b * (rugged_std - 0.215), 
         a ~ 0 + cid, 
         b ~ 0 + cid,
         nl = TRUE),
      prior = c(prior(normal(1, 0.1), class = b, coef = cid1, nlpar = a),
                prior(normal(1, 0.1), class = b, coef = cid2, nlpar = a),
                prior(normal(0, 0.3), class = b, coef = cid1, nlpar = b),
                prior(normal(0, 0.3), class = b, coef = cid2, nlpar = b),
                prior(exponential(1), class = sigma)),
      chains = 4, cores = 4,
      seed = 9)
```
```{r}
b9.1
```
#### The uniform prior should be dunif(0,10)

```{r}
b9.1_unif <- 
  brm(data = dd, 
      family = gaussian,
      bf(log_gdp_std ~ 0 + a + b * (rugged_std - 0.215), 
         a ~ 0 + cid, 
         b ~ 0 + cid,
         nl = TRUE),
      prior = c(prior(normal(1, 0.1), class = b, coef = cid1, nlpar = a),
                prior(normal(1, 0.1), class = b, coef = cid2, nlpar = a),
                prior(normal(0, 0.3), class = b, coef = cid1, nlpar = b),
                prior(normal(0, 0.3), class = b, coef = cid2, nlpar = b),
                prior(uniform(0,10), class = sigma)),
      chains = 4, cores = 4,
      seed = 9)
```
```{r}
b9.1_unif
```

> It looks like the different priors have no detectible influence on the posterior distribution.