---
title: "brms_week4_HW"
author: "Min-Yao"
date: "2020/9/10"
output: 
  html_document: 
    keep_md: yes
---

```{r}
library(rethinking)
library(brms)
library(tidyverse)
```

```{r}
germ <- read_csv("light_round1_tall.csv") %>%
  filter(wps == 0) %>%
  select(pops, temps, total_seeds, germ, day, cumulative_germ)
germ
```

#### Convert to one event per row
```{r}
one_per_row <- function(df) {
  total_seed <- max(df$total_seeds, sum(df$germ))
  newdata <- tibble(id=1:total_seed, germ=0, day=max(df$day))
  df <- df %>% filter(germ>0)
  count <- 1
  if (nrow(df) > 0) {
    for (i in 1:nrow(df)) { # we look at each row of the df where germination occured
      for (j in 1:df$germ[i]) { # now update the newdata to reflect the germiantion of each seed
        newdata$germ[count] <- 1
        newdata$day[count]=df$day[i]
        count <- count+1 # count keeps track of which individual we are at in the new data
      } # for j
    } # for i
  } # if 
  return(newdata)
}
germone <- germ %>% group_by(pops, temps) %>%
  select(-cumulative_germ) %>% # not needed in this encoding (I think...in any case would need to be recalculated)
  nest() %>%
  mutate(newdata=map(data, one_per_row)) %>%
  select(-data) %>%
  unnest(newdata)
germone
```
#### STDI only
```{r}
germ.stdi <- germone %>% filter(pops=="STDI") %>% select(-pops)
germ.stdi
```

# Assignment:

You can start from my models if you want to.  The three relevant ones from my code are

## m1.1/m1.2: censored exponential (the worst)

### m1.1: rethinking

exponential rate curve, censoring seed that don't germinate.

```{r}
d <- list(germ=germ.stdi$germ, 
          temps=as.numeric(as.factor(germ.stdi$temps)),
          day=germ.stdi$day)
m1.1 <- ulam(
  alist(
    day | germ==1 ~ exponential( lambda),
    day | germ==0 ~ custom(exponential_lccdf( !Y | lambda)),
    lambda <- 1.0 / mu,
    log(mu) <- a[temps],
    a[temps] ~ normal(0,1)),
  data=d,
  chains=4,
  cores = 4
)
```

```{r}
precis(m1.1, depth = 2)
```

The above represent log(mean time to germination)

### m1.2: brms censoring model

need to set up indicator for censoring.
```{r}
germ.stdi <- germ.stdi %>%
  mutate(cens=ifelse(germ==0, "right", "none"),
         tempsc=as.character(temps) %>% str_pad(width=2, pad="0"))
germ.stdi
```

```{r}
get_prior(day | cens(cens) ~ 0 + tempsc, family = exponential, data=germ.stdi)
```


```{r}
m1.2 <- brm(day | cens(cens) ~ 0 + tempsc,
            family = exponential(),
            set_prior("normal(0,1)", class="b"),
            data = germ.stdi, )
```

```{r}
summary(m1.2)
```

### m1.5: ZI exponential

Try to write a model with different dormancy probs for the different temps.
```{r}
d <- list(N=nrow(germ.stdi),
          germ=germ.stdi$germ,
          temps=as.numeric(as.factor(germ.stdi$temps)),
          day=germ.stdi$day)
stanmodel1.5 <-
  "
data{
    int<lower=1> N;  // number of observations
    int germ[N];
    vector[N] day;
    int temps[N];
}
parameters{
    vector[8] a; // alpha for the exponential curve, one for each temp
    vector[8] ap; // alpha for the proportion dormant, one for each temp
}
model{
    vector[N] p;
    vector[N] lambda;
    vector[N] mu;
    a ~ normal( 0 , 1 );
    ap ~ normal( 0, 1 );
    for (i in 1:N) {
        p[i] = ap[temps[i]];
        p[i] = inv_logit(p[i]); // inverse link function
    }
    for ( i in 1:N ) {
        mu[i] = a[temps[i]];
        mu[i] = exp(mu[i]); // inverse link function
    }
    for ( i in 1:N ) {
        lambda[i] = 1/mu[i];
    }
    for ( i in 1:N ) 
        if ( germ[i] == 0 ) target += log_mix(p[i], 0, exponential_lccdf(day[i] | lambda[i]));
    for ( i in 1:N ) 
        if ( germ[i] == 1 ) target += log1m(p[i]) +  exponential_lpdf(day[i] | lambda[i]);
    for ( i in 1:N ) 
        if ( germ[i] == 1 ) day[i] ~ exponential( lambda[i] );
}
"
m1.5 <- stan(model_code=stanmodel1.5, data=d)
```

```{r}
precis(m1.5, depth = 2)
```

### m1.7/m1.7a : ZI Gamma (the best)

Look at gamma stancode from brms:

```{r}
make_stancode(day ~ tempsc, 
             family="Gamma",
             prior=set_prior("normal(0,1)"),
             data=germ.stdi,
             sample=FALSE)
```


Try developing a ZI gamma model:

stan uses that gamma(alpha, beta) parameterization where "alpha" is shape" and "beta" is "rate".  So try having the shape be the same for all temps.
```{r}
d <- list(N=nrow(germ.stdi),
          germ=germ.stdi$germ,
          temps=as.numeric(as.factor(germ.stdi$temps)),
          day=germ.stdi$day)
stanmodel1.7 <-
  "
data{
    int<lower=1> N;  // number of observations
    int germ[N];
    vector[N] day;
    int temps[N];
}
parameters{
    real<lower=0> shape; // should set lower bound
    vector[8] a; // alpha for the gamma curve, one for each temp
    vector[8] ap; // alpha for the proportion dormant, one for each temp
}
model{
    vector[N] p;
    vector[N] mu;
    a ~ normal( 0 , .5 ); //narrow priors to overcome divergent transitions
    ap ~ normal( 0, 1.5 ); 
    shape ~ normal(0, .5); //narrow priors to overcome divergent transitions
    for (i in 1:N) {
        p[i] = ap[temps[i]];
        p[i] = inv_logit(p[i]);
    }
    for ( i in 1:N ) {
        mu[i] = a[temps[i]];
         // apply the inverse link function
        mu[i] = shape * exp(-(mu[i]));
    }
    for ( i in 1:N ) 
       if ( germ[i] == 0 ) target += log_mix(p[i], 0, gamma_lccdf(day[i] | shape, mu[i]));
    for ( i in 1:N ) 
       if ( germ[i] == 1 ) target += log1m(p[i]) + gamma_lpdf(day[i] | shape, mu[i]);
    for ( i in 1:N ) 
       if ( germ[i] == 1 ) day[i] ~ gamma( shape, mu[i] );
}
"
m1.7 <- stan(model_code=stanmodel1.7, data=d, chains=4, cores=4, control=list(adapt_delta=.99))
``` 


```{r}
precis(m1.7, depth = 2)
```


```{r}
#shinystan::launch_shinystan(m1.7)
```

playing with priors:

Make shape exponential

```{r}
d <- list(N=nrow(germ.stdi),
          germ=germ.stdi$germ,
          temps=as.numeric(as.factor(germ.stdi$temps)),
          day=germ.stdi$day)
stanmodel1.7a <-
  "
data{
    int<lower=1> N;  // number of observations
    int germ[N];
    vector[N] day;
    int temps[N];
}
parameters{
    real<lower=0> shape; // should set lower bound
    vector[8] a; // alpha for the gamma curve, one for each temp
    vector[8] ap; // alpha for the proportion dormant, one for each temp
}
model{
    vector[N] p;
    vector[N] mu;
    a ~ normal( 0 , .5 ); //narrow priors to overcome divergent transitions
    ap ~ normal( 0, 1.5 ); 
    shape ~ exponential(.5); 
    for (i in 1:N) {
        p[i] = ap[temps[i]];
        p[i] = inv_logit(p[i]);
    }
    for ( i in 1:N ) {
        mu[i] = a[temps[i]];
         // apply the inverse link function
        mu[i] = shape * exp(-(mu[i]));
    }
    for ( i in 1:N ) 
       if ( germ[i] == 0 ) target += log_mix(p[i], 0, gamma_lccdf(day[i] | shape, mu[i]));
    for ( i in 1:N ) 
       if ( germ[i] == 1 ) target += log1m(p[i]) + gamma_lpdf(day[i] | shape, mu[i]);
    for ( i in 1:N ) 
       if ( germ[i] == 1 ) day[i] ~ gamma( shape, mu[i] );
}
"
m1.7a <- stan(model_code=stanmodel1.7a, data=d, chains=4, cores=4, control=list(adapt_delta=.99))
``` 


```{r}
precis(m1.7a, depth = 2)
```

```{r}
post <- as.data.frame(m1.7a)
post_logavgdays <- post %>% select(shape, starts_with("a["
)) # these are the shape and the log(mean(avg time to germinate))
mu_rate <- post_logavgdays %>% 
  mutate(across(-shape, ~ shape*exp(-(.))) ) %>%
  summarize(across(everything(), mean)) %>% select(-shape) # posterior mean
mu_rate
mu_shape <- mean(post$shape)
post_logitp <- post %>% select(starts_with("ap")) #logit p dormant
mu_p <- post_logitp %>%
  summarize_all(mean) %>%
  mutate_all(inv_logit)
mu_p
posterior_coef <- tibble(shape=mu_shape, temps=as.factor(unique(germ.stdi$temps)),
                         rate=t(mu_rate),
                         p=t(mu_p))
post_plot1.7a <- expand_grid(posterior_coef, day=1:28) %>%
  mutate(prop_germ=pgamma(day, shape=mu_shape, rate=rate) * (1-p))
stdi.plot <- germ %>% filter(pops=="STDI") %>% 
  select(day, temps, cumulative_germ, total_seeds) %>%
  mutate(temps=as.factor(temps),
         prop_germ=cumulative_germ/total_seeds)
post_plot1.7a %>% 
  ggplot(aes(x=day,y=prop_germ,color=temps,group=temps)) +
  geom_line() +
  geom_point(data=stdi.plot)
```

### Choose at least one model from above (or your own) to work from. 
### From the base model:

### 1) Try using temperature as a continuous predictor.  Perhaps as a quadratic, or perhaps using the equation that Rongkui showed. (Or try both).

from m1.7a : ZI Gamma (the best)

```{r}
d <- list(N=nrow(germ.stdi),
          germ=germ.stdi$germ,
          temps=as.vector(scale(germ.stdi$temps)),
          day=germ.stdi$day)

stanmodel1.7a_quad <-
  "
data{
    int<lower=1> N;  // number of observations
    int germ[N];
    vector[N] day;
    real temps[N];
}
parameters{
    real<lower=0> shape; // should set lower bound
    real a; // alpha for the gamma curve, one for each temp
    real ap; // alpha for the proportion dormant, one for each temp
    real b; // b for the gamma curve, one for each temp
    real bp; // b for the proportion dormant, one for each temp
    real c; // c for the gamma curve, one for each temp
    real cp; // c for the proportion dormant, one for each temp
}
model{
    vector[N] p;
    vector[N] mu;
    a ~ normal( 0 , .5 ); //narrow priors to overcome divergent transitions
    ap ~ normal( 0, 1.5 );
    b ~ normal( 0 , .5 ); //narrow priors to overcome divergent transitions
    bp ~ normal( 0, 1.5 );
    c ~ normal( 0 , .5 ); //narrow priors to overcome divergent transitions
    cp ~ normal( 0, 1.5 );
    shape ~ exponential(.5); 
    for (i in 1:N) {
        p[i] = ap + bp*temps[i] + cp*(temps[i]^2);
        p[i] = inv_logit(p[i]);
    }
    for ( i in 1:N ) {
        mu[i] = a + b*temps[i] + c*(temps[i]^2);
         // apply the inverse link function
        mu[i] = shape * exp(-(mu[i]));
    }
    for ( i in 1:N ) 
       if ( germ[i] == 0 ) target += log_mix(p[i], 0, gamma_lccdf(day[i] | shape, mu[i]));
    for ( i in 1:N ) 
       if ( germ[i] == 1 ) target += log1m(p[i]) + gamma_lpdf(day[i] | shape, mu[i]);
    for ( i in 1:N ) 
       if ( germ[i] == 1 ) day[i] ~ gamma( shape, mu[i] );
}
"
m1.7a_quad <- stan(model_code=stanmodel1.7a_quad, data=d, chains=4, cores=4, control=list(adapt_delta=.99))
``` 

```{r}
precis(m1.7a_quad)
```

```{r}
post <- as.data.frame(m1.7a_quad)
predict_post <- function(temp, post, days=1:28) {
    p <- post$ap + post$bp*temp + post$bpq*temp^2
    p <- inv_logit(p)
    mu <- post$a + post$bg*temp + post$bgq*temp^2
    mu <- post$shape*exp(-mu)
    postparams <- tibble(id = 1:length(p), p=p, mu=mu, shape=post$shape) %>%
      nest(params=-id)
    prop_germ <- postparams %>% 
      mutate(prop_germ=map(params, ~ pgamma(days, shape=.$shape, rate=.$mu) * (1-.$p))) %>% 
      select(-params) %>% 
      unnest(prop_germ) %>%
      mutate(day=rep(days, length.out=nrow(.)))
}
post_predictions <- tibble(realtemp = unique(germ.stdi$temps)) %>%
  mutate(scaletemp = scale(realtemp)) %>%
  mutate(predictions = map(scaletemp, predict_post, post)) %>% unnest(predictions)
post_predictions_summary <- post_predictions %>%
  group_by(realtemp, day) %>%
  summarize(low89=HPDI(prop_germ)[1], 
            high90=HPDI(prop_germ)[2],
            prop_germ=mean(prop_germ)) %>%
  mutate(temps=as.factor(realtemp))
stdi.plot <- germ %>% filter(pops=="STDI") %>% 
  select(day, temps, cumulative_germ, total_seeds) %>%
  mutate(temps=as.factor(temps),
         prop_germ=cumulative_germ/total_seeds)
post_predictions_summary %>% 
  ggplot(aes(x=day,y=prop_germ,color=temps,group=temps)) +
  geom_line() +
  geom_point(data=stdi.plot)
```


from m1.5: ZI exponential

```{r}
d <- list(N=nrow(germ.stdi),
          germ=germ.stdi$germ,
          temps=as.numeric(scale(germ.stdi$temps)),
          day=germ.stdi$day)

stanmodel1.5_quad <-
  "
data{
    int<lower=1> N;  // number of observations
    int germ[N];
    vector[N] day;
    int temps[N];
}
parameters{
    real a; // alpha for the exponential curve, one for each temp
    real ap; // alpha for the proportion dormant, one for each temp
    real b; // b for the exponential curve, one for each temp
    real bp; // b for the proportion dormant, one for each temp
    real c; // c for the exponential curve, one for each temp
    real cp; // c for the proportion dormant, one for each temp
}
model{
    vector[N] p;
    vector[N] lambda;
    vector[N] mu;
    a ~ normal( 0 , 1 );
    ap ~ normal( 0, 1 );
    b ~ normal( 0 , .5 ); //narrow priors to overcome divergent transitions
    bp ~ normal( 0, 1.5 );
    c ~ normal( 0 , .5 ); //narrow priors to overcome divergent transitions
    cp ~ normal( 0, 1.5 );
    for (i in 1:N) {
        p[i] = ap + bp*temps[i] + cp*(temps[i]^2);
        p[i] = inv_logit(p[i]); // inverse link function
    }
    for ( i in 1:N ) {
        mu[i] = a + b*temps[i] + c*(temps[i]^2);
        mu[i] = exp(mu[i]); // inverse link function
    }
    for ( i in 1:N ) {
        lambda[i] = 1/mu[i];
    }
    for ( i in 1:N ) 
        if ( germ[i] == 0 ) target += log_mix(p[i], 0, exponential_lccdf(day[i] | lambda[i]));
    for ( i in 1:N ) 
        if ( germ[i] == 1 ) target += log1m(p[i]) +  exponential_lpdf(day[i] | lambda[i]);
    for ( i in 1:N ) 
        if ( germ[i] == 1 ) day[i] ~ exponential( lambda[i] );
}
"
m1.5_quad <- stan(model_code=stanmodel1.5_quad, data=d)
```

```{r}
precis(m1.5_quad, depth = 2)
```



### 2) Try incorporating multiple populations.

```{r}
d <- list(N=nrow(germone),
          germ=germone$germ,
          temps=as.numeric(as.factor(germone$temps)),
          pops=as.numeric(as.factor(germone$pops)),
          day=germone$day)

str(d)

stanmodel1.7_m <-
  "
data{
    int<lower=1> N;  // number of observations
    int germ[N];
    vector[N] day;
    int temps[N];
    int pops[N];
}
parameters{
    real<lower=0> shape; // should set lower bound
    vector[8] a; // alpha for the gamma curve, one for each temp
    vector[8] ap; // alpha for the proportion dormant, one for each temp
    vector[24] b; // b for the gamma curve, one for each pop
    vector[24] bp; // b for the proportion dormant, one for each pop
}
model{
    vector[N] p;
    vector[N] mu;
    a ~ normal( 0 , .5 ); //narrow priors to overcome divergent transitions
    ap ~ normal( 0, 1.5 ); 
    b ~ normal( 0 , .5 ); //narrow priors to overcome divergent transitions
    bp ~ normal( 0, 1.5 ); 
    shape ~ exponential(.5); 
    for (i in 1:N) {
        p[i] = ap[temps[i]]+bp[pops[i]];
        p[i] = inv_logit(p[i]);
    }
    for ( i in 1:N ) {
        mu[i] = a[temps[i]]+b[pops[i]];
         // apply the inverse link function
        mu[i] = shape * exp(-(mu[i]));
    }
    for ( i in 1:N ) 
       if ( germ[i] == 0 ) target += log_mix(p[i], 0, gamma_lccdf(day[i] | shape, mu[i]));
    for ( i in 1:N ) 
       if ( germ[i] == 1 ) target += log1m(p[i]) + gamma_lpdf(day[i] | shape, mu[i]);
    for ( i in 1:N ) 
       if ( germ[i] == 1 ) day[i] ~ gamma( shape, mu[i] );
}
"
m1.7_m <- stan(model_code=stanmodel1.7_m, data=d, chains=4, cores=4, control=list(adapt_delta=.99))
``` 

```{r}
precis(m1.7_m, depth = 2)
```

```{r}
post <- as.data.frame(m1.7_m)
post_logavgdays <- post %>% select(shape, starts_with("a["
)) # these are the shape and the log(mean(avg time to germinate))
mu_rate <- post_logavgdays %>% 
  mutate(across(-shape, ~ shape*exp(-(.))) ) %>%
  summarize(across(everything(), mean)) %>% select(-shape) # posterior mean
mu_rate
mu_shape <- mean(post$shape)
post_logitp <- post %>% select(starts_with("ap")) #logit p dormant
mu_p <- post_logitp %>%
  summarize_all(mean) %>%
  mutate_all(inv_logit)
mu_p
posterior_coef <- tibble(shape=mu_shape, temps=as.factor(unique(germone$temps)),
                         rate=t(mu_rate),
                         p=t(mu_p))
post_plot1.7_m <- expand_grid(posterior_coef, day=1:28) %>%
  mutate(prop_germ=pgamma(day, shape=mu_shape, rate=rate) * (1-p))
stdi.plot <- germ %>% filter(pops=="STDI") %>% 
  select(day, temps, cumulative_germ, total_seeds) %>%
  mutate(temps=as.factor(temps),
         prop_germ=cumulative_germ/total_seeds)
post_plot1.7_m %>% 
  ggplot(aes(x=day,y=prop_germ,color=temps,group=temps)) +
  geom_line() +
  geom_point(data=stdi.plot)
```

