---
title: "chapter 5"
author: "Min-Yao"
date: "2019年5月8日"
output: 
  html_document: 
    keep_md: yes
---

# The Many Variables & The Spurious Waffles
## 5.1. Spurious association

```{r}
## R code 5.1
# load data and copy
library(rethinking)
data(WaffleDivorce)
d <- WaffleDivorce

# standardize variables
d$A <- scale( d$MedianAgeMarriage )
d$D <- scale( d$Divorce )
```

```{r}
## R code 5.2
sd( d$MedianAgeMarriage )
```

```{r}
## R code 5.3
m5.1 <- quap(
    alist(
        D ~ dnorm( mu , sigma ) ,
        mu <- a + bA * A ,
        a ~ dnorm( 0 , 0.2 ) ,
        bA ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
    ) , data = d )
```

```{r}
## R code 5.4
set.seed(10)
prior <- extract.prior( m5.1 )
mu <- link( m5.1 , post=prior , data=list( A=c(-2,2) ) )
plot( NULL , xlim=c(-2,2) , ylim=c(-2,2) )
for ( i in 1:50 ) lines( c(-2,2) , mu[i,] , col=col.alpha("black",0.4) )
```


```{r}
## R code 5.5
# compute percentile interval of mean
A_seq <- seq( from=-3 , to=3.2 , length.out=30 )
mu <- link( m5.1 , data=list(A=A_seq) )
mu.mean <- apply( mu , 2, mean )
mu.PI <- apply( mu , 2 , PI )

# plot it all
plot( D ~ A , data=d , col=rangi2 )
lines( A_seq , mu.mean , lwd=2 )
shade( mu.PI , A_seq )
```

```{r}
## R code 5.6
d$M <- scale( d$Marriage )
m5.2 <- quap(
    alist(
        D ~ dnorm( mu , sigma ) ,
        mu <- a + bM * M ,
        a ~ dnorm( 0 , 0.2 ) ,
        bM ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
    ) , data = d )
```

```{r}
## R code 5.4
set.seed(10)
prior <- extract.prior( m5.2 )
mu <- link( m5.2 , post=prior , data=list( M=c(-2,2) ) )
plot( NULL , xlim=c(-2,2) , ylim=c(-2,2) )
for ( i in 1:50 ) lines( c(-2,2) , mu[i,] , col=col.alpha("black",0.4) )
```

```{r}
## R code 5.5
# compute percentile interval of mean
M_seq <- seq( from=-3 , to=3.2 , length.out=30 )
mu <- link( m5.2 , data=list(M=M_seq) )
mu.mean <- apply( mu , 2, mean )
mu.PI <- apply( mu , 2 , PI )

# plot it all
plot( D ~ M , data=d , col=rangi2 )
lines( M_seq , mu.mean , lwd=2 )
shade( mu.PI , M_seq )
```

### 5.1.1. Think before you regress.

```{r}
## R code 5.7
#install.packages('dagitty')
library(dagitty)
dag5.1 <- dagitty( "dag {
    A -> D
    A -> M
    M -> D
}")
coordinates(dag5.1) <- list( x=c(A=0,D=1,M=2) , y=c(A=0,D=1,M=0) )
plot( dag5.1 )
```

### 5.1.2. Multiple regression notation. 

```{r}
## R code 5.8
m5.3 <- quap(
    alist(
        D ~ dnorm( mu , sigma ) ,
        mu <- a + bM*M + bA*A ,
        a ~ dnorm( 0 , 0.2 ) ,
        bM ~ dnorm( 0 , 0.5 ) ,
        bA ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
    ) , data = d )
precis( m5.3 )

## R code 5.9
plot(coeftab(m5.1,m5.2,m5.3), par=c("bA","bM") )
```


```{r}
## R code 5.10
N <- 50 # number of simulated States
age <- rnorm( N )        # sim A
mar <- rnorm( N , age )  # sim A -> M
div <- rnorm( N , age )  # sim A -> D
```

### 5.1.4. Plotting multivariate posteriors.

```{r}
## R code 5.11
m5.4 <- quap(
    alist(
        M ~ dnorm( mu , sigma ) ,
        mu <- a + bAM * A ,
        a ~ dnorm( 0 , 0.2 ) ,
        bAM ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
    ) , data = d )
```

```{r}
## R code 5.12
mu <- link(m5.4)
mu_mean <- apply( mu , 2 , mean )
mu_resid <- d$M - mu_mean
```


```{r}
## R code 5.13
# prepare new counterfactual data
M_seq <- seq( from=-2 , to=3 , length.out=30 )
pred_data <- data.frame( M = M_seq , A = 0 )

# compute counterfactual mean divorce (mu)
mu <- link( m5.3 , data=pred_data )
mu_mean <- apply( mu , 2 , mean )
mu_PI <- apply( mu , 2 , PI )

# simulate counterfactual divorce outcomes
D_sim <- sim( m5.3 , data=pred_data , n=1e4 )
D_PI <- apply( D_sim , 2 , PI )

# display predictions, hiding raw data with type="n"
plot( D ~ M , data=d , type="n" )
mtext( "Median age marriage (std) = 0" )
lines( M_seq , mu_mean )
shade( mu_PI , M_seq )
shade( D_PI , M_seq )
```

```{r}
## R code 5.13
# prepare new counterfactual data
A_seq <- seq( from=-3 , to=4 , length.out=30 )
pred_data <- data.frame( A = A_seq , M = 0 )

# compute counterfactual mean divorce (mu)
mu <- link( m5.3 , data=pred_data )
mu_mean <- apply( mu , 2 , mean )
mu_PI <- apply( mu , 2 , PI )

# simulate counterfactual divorce outcomes
D_sim <- sim( m5.3 , data=pred_data , n=1e4 )
D_PI <- apply( D_sim , 2 , PI )

# display predictions, hiding raw data with type="n"
plot( D ~ A , data=d , type="n" )
mtext( "Median age marriage (std) = 0" )
lines( A_seq , mu_mean )
shade( mu_PI , A_seq )
shade( D_PI , A_seq )
```

```{r}
## R code 5.14
# call link without specifying new data
# so it uses original data
mu <- link( m5.3 )

# summarize samples across cases
mu_mean <- apply( mu , 2 , mean )
mu_PI <- apply( mu , 2 , PI )

# simulate observations
# again no new data, so uses original data
D_sim <- sim( m5.3 , n=1e4 )
D_PI <- apply( D_sim , 2 , PI )
```

```{r}
## R code 5.15
plot( mu_mean ~ d$D , col=rangi2 , ylim=range(mu_PI) ,
    xlab="Observed divorce" , ylab="Predicted divorce" )
abline( a=0 , b=1 , lty=2 )
for ( i in 1:nrow(d) ) lines( rep(d$D[i],2) , mu_PI[,i] , col=rangi2 )

## R code 5.16
identify( x=d$D , y=mu_mean , labels=d$Loc )
```


```{r}
## R code 5.17
N <- 100                         # number of cases
x_real <- rnorm( N )             # x_real as Gaussian with mean 0 and stddev 1
x_spur <- rnorm( N , x_real )    # x_spur as Gaussian with mean=x_real
y <- rnorm( N , x_real )         # y as Gaussian with mean=x_real
d <- data.frame(y,x_real,x_spur) # bind all together in data frame
```

## 5.2. Masked relationship

```{r}
## R code 5.18
library(rethinking)
data(milk)
d <- milk
str(d)
```


```{r}
## R code 5.19
d$K <- scale( d$kcal.per.g )
d$N <- scale( d$neocortex.perc )
d$M <- scale( log(d$mass) )
```

```{r}
## R code 5.20
#m5.5_draft <- quap(
#    alist(
#        K ~ dnorm( mu , sigma ) ,
#        mu <- a + bN*N ,
#        a ~ dnorm( 0 , 1 ) ,
#        bN ~ dnorm( 0 , 1 ) ,
#        sigma ~ dexp( 1 )
#    ) , data=d )
```


```{r}
## R code 5.21
d$neocortex.perc
```

```{r}
## R code 5.22
dcc <- d[ complete.cases(d$K,d$N,d$M) , ]
```

```{r}
## R code 5.23
m5.5_draft <- quap(
    alist(
        K ~ dnorm( mu , sigma ) ,
        mu <- a + bN*N ,
        a ~ dnorm( 0 , 1 ) ,
        bN ~ dnorm( 0 , 1 ) ,
        sigma ~ dexp( 1 )
    ) , data=dcc )
```


```{r}
## R code 5.24
prior <- extract.prior( m5.5_draft )
xseq <- c(-2,2)
mu <- link( m5.5_draft , post=prior , data=list(N=xseq) )
plot( NULL , xlim=xseq , ylim=xseq )
for ( i in 1:50 ) lines( xseq , mu[i,] , col=col.alpha("black",0.3) )
```


```{r}
## R code 5.25
m5.5 <- quap(
    alist(
        K ~ dnorm( mu , sigma ) ,
        mu <- a + bN*N ,
        a ~ dnorm( 0 , 0.2 ) ,
        bN ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
    ) , data=dcc )

## R code 5.26
precis( m5.5 )
```
```{r}
## R code 5.24
prior <- extract.prior( m5.5 )
xseq <- c(-2,2)
mu <- link( m5.5 , post=prior , data=list(N=xseq) )
plot( NULL , xlim=xseq , ylim=xseq )
for ( i in 1:50 ) lines( xseq , mu[i,] , col=col.alpha("black",0.3) )
```

```{r}
## R code 5.27
xseq <- seq( from=min(dcc$N)-0.15 , to=max(dcc$N)+0.15 , length.out=30 )
mu <- link( m5.5 , data=list(N=xseq) )
mu_mean <- apply(mu,2,mean)
mu_PI <- apply(mu,2,PI)
plot( K ~ N , data=dcc )
lines( xseq , mu_mean , lwd=2 )
shade( mu_PI , xseq )
```

```{r}
## R code 5.28
m5.6 <- quap(
    alist(
        K ~ dnorm( mu , sigma ) ,
        mu <- a + bM*M ,
        a ~ dnorm( 0 , 0.2 ) ,
        bM ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
    ) , data=dcc )
precis(m5.6)
```

```{r}
## R code 5.29
m5.7 <- quap(
    alist(
        K ~ dnorm( mu , sigma ) ,
        mu <- a + bN*N + bM*M ,
        a ~ dnorm( 0 , 0.2 ) ,
        bN ~ dnorm( 0 , 0.5 ) ,
        bM ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
    ) , data=dcc )
precis(m5.7)

## R code 5.30
plot( coeftab( m5.5 , m5.6 , m5.7 ) , pars=c("bM","bN") )
```

```{r}
## R code 5.31
xseq <- seq( from=min(dcc$M)-0.15 , to=max(dcc$M)+0.15 , length.out=30 )
mu <- link( m5.7 , data=data.frame( M=xseq , N=0 ) )
mu_mean <- apply(mu,2,mean)
mu_PI <- apply(mu,2,PI)
plot( NULL , xlim=range(dcc$M) , ylim=range(dcc$K) )
lines( xseq , mu_mean , lwd=2 )
shade( mu_PI , xseq )
```

```{r}
## R code 5.32
# M -> K <- N
# M -> N
n <- 100
M <- rnorm( n )
N <- rnorm( n , M )
K <- rnorm( n , N - M )
d_sim <- data.frame(K=K,N=N,M=M)
```

```{r}
## R code 5.29
m5.7sim1 <- quap(
    alist(
        K ~ dnorm( mu , sigma ) ,
        mu <- a + bN*N + bM*M ,
        a ~ dnorm( 0 , 0.2 ) ,
        bN ~ dnorm( 0 , 0.5 ) ,
        bM ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
    ) , data=d_sim )
precis(m5.7sim1)
```
```{r}
## R code 5.33
# M -> K <- N
# N -> M
n <- 100
N <- rnorm( n )
M <- rnorm( n , N )
K <- rnorm( n , N - M )
d_sim2 <- data.frame(K=K,N=N,M=M)
```

```{r}
## R code 5.29
m5.7sim2 <- quap(
    alist(
        K ~ dnorm( mu , sigma ) ,
        mu <- a + bN*N + bM*M ,
        a ~ dnorm( 0 , 0.2 ) ,
        bN ~ dnorm( 0 , 0.5 ) ,
        bM ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
    ) , data=d_sim2 )
precis(m5.7sim2)
```

```{r}
# M -> K <- N
# M <- U -> N
n <- 100
U <- rnorm( n )
N <- rnorm( n , U )
M <- rnorm( n , U )
K <- rnorm( n , N - M )
d_sim3 <- data.frame(K=K,N=N,M=M)
```

```{r}
## R code 5.29
m5.7sim3 <- quap(
    alist(
        K ~ dnorm( mu , sigma ) ,
        mu <- a + bN*N + bM*M ,
        a ~ dnorm( 0 , 0.2 ) ,
        bN ~ dnorm( 0 , 0.5 ) ,
        bM ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
    ) , data=d_sim3 )
precis(m5.7sim3)
```

## 5.3. Categorical variables

```{r}
## R code 5.34
data(Howell1)
d <- Howell1
str(d)
```

```{r}
## R code 5.35
mu_female <- rnorm(1e4,178,20)
mu_male <- rnorm(1e4,178,20) + rnorm(1e4,0,10)
precis( data.frame( mu_female , mu_male ) )
```


```{r}
## R code 5.36
d$sex <- ifelse( d$male==1 , 2 , 1 )
str( d$sex )
```


```{r}
## R code 5.37
m5.8 <- quap(
    alist(
        height ~ dnorm( mu , sigma ) ,
        mu <- a[sex] ,
        a[sex] ~ dnorm( 178 , 20 ) ,
        sigma ~ dunif( 0 , 50 )
    ) , data=d )
precis( m5.8 , depth=2 )
```


```{r}
## R code 5.38
post <- extract.samples(m5.8)
post$diff_fm <- post$a[,1] - post$a[,2]
precis( post , depth=2 )
```

```{r}
## R code 5.39
data(milk)
d <- milk
unique(d$clade)
```

```{r}
## R code 5.40
d$clade_id <- as.integer( d$clade )
```

```{r}
## R code 5.41
d$K <- scale( d$kcal.per.g )
m5.9 <- quap(
    alist(
        K ~ dnorm( mu , sigma ),
        mu <- a[clade_id],
        a[clade_id] ~ dnorm( 0 , 0.5 ),
        sigma ~ dexp( 1 )
    ) , data=d )
labels <- paste( "a[" , 1:4 , "]:" , levels(d$clade) , sep="" )
plot( precis( m5.9 , depth=2 , pars="a" ) , labels=labels , xlab="expected kcal (std)" )
```

```{r}
## R code 5.42
set.seed(63)
d$house <- sample( rep(1:4,each=8) , size=nrow(d) )

## R code 5.43
m5.10 <- quap(
    alist(
        K ~ dnorm( mu , sigma ),
        mu <- a[clade_id] + h[house],
        a[clade_id] ~ dnorm( 0 , 0.5 ),
        h[house] ~ dnorm( 0 , 0.5 ),
        sigma ~ dexp( 1 )
    ) , data=d )
```


## Easy.
### 5E1. Which of the linear models below are multiple linear regressions?
(1) 
$$
\mu_i = \alpha + \beta x_i
$$
(2) 
$$
\mu_i = \beta_x x_i + \beta_z z_i
$$
(3) 
$$
\mu_i = \alpha + \beta (x_i-z_i)
$$
(4) 
$$
\mu_i = \alpha + \beta_x x_i + \beta_z z_i
$$

>  2, 3, and 4 are multiple regressions because these models have two predictors.

### 5E3.  Write down a multiple regression to evaluate the claim: Neither amount of funding nor size of laboratory is by itself a good predictor of time to PhD degree; but together these variables are both positively associated with time to degree. Write down the model definition and indicate which side of zero each slope parameter should be on.

> F is amount of funding; S is size of laboratory.

$$
\mu_i = \alpha + \beta_F F_i + \beta_S S_i
$$
$\beta_F$ and $\beta_S$ should be positive.

## Medium.
### 5M2. Invent your own example of a masked relationship. An outcome variable should be correlated with both predictor variables, but in opposite directions. And the two predictor variables should be correlated with one another.

```{r}
## change from R code 5.17
library(ggplot2)
library(GGally)
set.seed(1)
N <- 1000                         # number of cases
x_real <- rnorm( N )             # x_real as Gaussian with mean 0 and stddev 1
x_spur <- rnorm( N , x_real*0.5 )    # x_spur as Gaussian with mean=x_real*0.5
y <- rnorm( N , x_real-x_spur )         # y as Gaussian with mean=x_real-x_spur
d <- data.frame(y,x_real,x_spur) # bind all together in data frame
ggpairs(d)
precis(lm(y~x_spur, d))
precis(lm(y~x_real, d))
precis(lm(y~x_real+x_spur, d))
```

### 5M3. It is sometimes observed that the best predictor of fire risk is the presence of firefighters—States and localities with many firefighters also have more fires. Presumably firefighters do not cause fires. Nevertheless, this is not a spurious correlation. Instead fires cause firefighters. Consider the same reversal of causal inference in the context of the divorce and marriage data. How might a high divorce rate cause a higher marriage rate? Can you think of a way to evaluate this relationship, using multiple regression?

> I think a high divorce rate may cause more single individuals and these divorced people can also marriage again. Therefore, to evaluate this relationship, I can use multiple regression by regressing marriage rate on both divorce rate and non-first marriage rate. If there is little or no additional predictive power in also knowing divorce rate once we know non-first marriage rate, this means my hypothesis is correct.


### 5M4. In the divorce data, States with high numbers of Mormons (members of The Church of Jesus Christ of Latter-day Saints, LDS) have much lower divorce rates than the regression models expected. Find a list of LDS population by State and use those numbers as a predictor variable, predicting divorce rate using marriage rate, median age at marriage, and percent LDS population (possibly standardized). You may want to consider transformations of the raw percent LDS variable.

```{r}
library(rethinking)
data(WaffleDivorce)
d <- WaffleDivorce
str(d)
d$Location
```

> Download data from https://www.worldatlas.com/articles/mormon-population-by-state.html and save it as CSV


```{r}
LDS <- read.csv("LDS.csv", header = TRUE)
LDS
str(LDS)
```

```{r}
d$LDS <- LDS$PercentageofMormonResidents
str(d)
```

```{r}
d$logLDS <- log(d$LDS)
d$L <- scale(d$logLDS)
d$A <- scale(d$MedianAgeMarriage)
d$D <- scale(d$Divorce)
d$M <- scale(d$Marriage)
str(d)
```

```{r}
## change from R code 5.8
m5M4 <- quap(
    alist(
        D ~ dnorm( mu , sigma ) ,
        mu <- a + bM*M + bA*A + bL*L ,
        a ~ dnorm( 0 , 0.2 ) ,
        bM ~ dnorm( 0 , 0.5 ) ,
        bA ~ dnorm( 0 , 0.5 ) ,
        bL ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
    ) , data = d )
precis( m5M4 )

## R code 5.9
plot(coeftab(m5.1,m5.2,m5.3,m5M4), par=c("bA","bM","bL") )
```

## Hard. 
### All three exercises below use the same data, data(foxes) (part of rethinking). 79 The urban fox (Vulpes vulpes) is a successful exploiter of human habitat. Since urban foxes move in packs and defend territories, data on habitat quality and population density is also included. The data frame has five columns:
(1) group: Number of the social group the individual fox belongs to
(2) avgfood: Th e average amount of food available in the territory
(3) groupsize: Th e number of foxes in the social group
(4) area: Size of the territory
(5) weight: Body weight of the individual fox

### 5H1. Fit two bivariate Gaussian regressions, using quap: (1) body weight as a linear function of territory size (area), and (2) body weight as a linear function of groupsize. Plot the results of these regressions, displaying the MAP regression line and the 95% interval of the mean. Is either variable important for predicting fox body weight?

```{r}
library(rethinking)
data(foxes)
d <- foxes
str(d)
```

```{r}
d$W <- scale(d$weight)
d$A <- scale(d$area)
d$S <- scale(d$groupsize)
str(d)
```


```{r}
## R code 5.3
m5H1.1 <- quap(
    alist(
        W ~ dnorm( mu , sigma ) ,
        mu <- a + bA * A ,
        a ~ dnorm( 0 , 0.2 ) ,
        bA ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
    ) , data = d )
precis(m5H1.1)
```

```{r}
## R code 5.5
# compute percentile interval of mean
A_seq <- seq( from= min(d$A) , to= max(d$A) , length.out=30 )
mu <- link( m5H1.1 , data=list(A=A_seq) )
mu.mean <- apply( mu , 2, mean )
mu.PI <- apply( mu , 2 , PI )

# plot it all
plot( W ~ A , data=d , col=rangi2 )
lines( A_seq , mu.mean , lwd=2 )
shade( mu.PI , A_seq )
```

```{r}
## R code 5.3
m5H1.2 <- quap(
    alist(
        W ~ dnorm( mu , sigma ) ,
        mu <- a + bS * S ,
        a ~ dnorm( 0 , 0.2 ) ,
        bS ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
    ) , data = d )
precis(m5H1.2)
```

```{r}
## R code 5.5
# compute percentile interval of mean
S_seq <- seq( from= min(d$S) , to= max(d$S) , length.out=30 )
mu <- link( m5H1.2 , data=list(S=S_seq) )
mu.mean <- apply( mu , 2, mean )
mu.PI <- apply( mu , 2 , PI )

# plot it all
plot( W ~ S , data=d , col=rangi2 )
lines( S_seq , mu.mean , lwd=2 )
shade( mu.PI , S_seq )
```


### 5H2. Now fit a multiple linear regression with weight as the outcome and both area and groupsize as predictor variables. Plot the predictions of the model for each predictor, holding the other predictor constant at its mean. What does this model say about the importance of each variable? Why do you get diff erent results than you got in the exercise just above?

```{r}
## R code 5.8
m5H2 <- quap(
    alist(
        W ~ dnorm( mu , sigma ) ,
        mu <- a + bS*S + bA*A ,
        a ~ dnorm( 0 , 0.2 ) ,
        bS ~ dnorm( 0 , 0.5 ) ,
        bA ~ dnorm( 0 , 0.5 ) ,
        sigma ~ dexp( 1 )
    ) , data = d )
precis( m5H2 )

## R code 5.9
#plot(coeftab(m5H1.1,m5H1.2,m5H2), par=c("bS","bA") )
```


```{r}
## R code 5.13
# prepare new counterfactual data
S_seq <- seq( from= min(d$S) , to= max(d$S) , length.out=30 )
pred_data <- data.frame( S = S_seq , A = 0 )

# compute counterfactual mean divorce (mu)
mu <- link( m5H2 , data=pred_data )
mu_mean <- apply( mu , 2 , mean )
mu_PI <- apply( mu , 2 , PI )

# simulate counterfactual divorce outcomes
W_sim <- sim( m5H2 , data=pred_data , n=1e4 )
W_PI <- apply( W_sim , 2 , PI )

# display predictions, hiding raw data with type="n"
plot( W ~ S , data=d , type="n" )
mtext( "Area (std) = 0" )
lines( S_seq , mu_mean )
shade( mu_PI , S_seq )
shade( W_PI , S_seq )
```

```{r}
## R code 5.13
# prepare new counterfactual data
A_seq <- seq( from= min(d$A) , to= max(d$A) , length.out=30 )
pred_data <- data.frame( A = A_seq , S = 0 )

# compute counterfactual mean divorce (mu)
mu <- link( m5H2 , data=pred_data )
mu_mean <- apply( mu , 2 , mean )
mu_PI <- apply( mu , 2 , PI )

# simulate counterfactual divorce outcomes
W_sim <- sim( m5H2 , data=pred_data , n=1e4 )
W_PI <- apply( W_sim , 2 , PI )

# display predictions, hiding raw data with type="n"
plot( W ~ A , data=d , type="n" )
mtext( "Groupsize (std) = 0" )
lines( A_seq , mu_mean )
shade( mu_PI , A_seq )
shade( W_PI , A_seq )
```