---
title: "HW_01_30_2020"
author: "Min-Yao"
date: "2020/1/30"
output: 
  html_document: 
    keep_md: yes
---

This csv has measurements of tomato internodes and petioles from wildtype (Moneymaker) and various phytochrome mutant lines.  Measurements were made at 3 time points, 21, 28, and 35 days after germination under two treatments, simulated sun and simulated shade.

for today let's focus on day 35.  Also let's focus on total stem length.  So: first


```{r}
phyE <- read.csv("figure4phyE.csv")
head(phyE)
summary(phyE)
str(phyE)
```

### Q1)

#### a) subset the data for day 35

```{r}
newphyE <- subset(phyE, day == 35)
head(newphyE)
summary(newphyE)
str(newphyE)
```

#### b) create a new column "stem_length" that is the sum of epi, int1, int2, and int3

```{r}
library(tidyverse)
newphyE_S <- newphyE %>% mutate(stem_length = rowSums(.[5:8]))
head(newphyE_S)
summary(newphyE_S)
str(newphyE_S)

```



#### c) although flats are listed as 1-6, flats in sun and shade are separate. Create a new column "flat2" that corrects for this.

```{r}
newphyE_S_flat2 <- newphyE_S %>% unite("flat2", c("treatment", "flat"), remove = FALSE)
head(newphyE_S_flat2)
summary(newphyE_S_flat2)
str(newphyE_S_flat2)
```


#### Ultimately you want to know if any of the mutants have a different length from Moneymaker, in sun or in shade, or if the response to shade differs.

```{r}
library(rethinking)

newphyE_list <- list(
  genotype = as.integer(newphyE_S_flat2$genotype),
  treatment_s = as.integer(newphyE_S_flat2$treatment=="shade"),
  flat_factor = as.integer(as.factor(newphyE_S_flat2$flat2)),
  stem_length = newphyE_S_flat2$stem_length
)

head(newphyE_list)
summary(newphyE_list)
str(newphyE_list)
```


### Q2) Fit 3 models, all of which include genotype and treatment
#### a) don't include flat.  Determine whether genotype, treatment, and their interaction are important predictors of stem_length

```{r}
modle_base <- ulam(
  alist(
    stem_length ~ dnorm(mu, sigma),
    mu <- a[genotype] + bT*treatment_s + bI[genotype]*treatment_s,
    a[genotype] ~ dnorm(140, 30),
    bT ~ dnorm(0, 30),
    bI[genotype] ~ dnorm(0, 20),
    sigma ~ dexp(1)
  ), data=newphyE_list , chains=4 , cores=4 , log_lik=TRUE)

precis( modle_base, depth=2 )
plot( precis(modle_base,depth=2) )
```


#### b) include flat without pooling

```{r}
modle_fixed <- ulam(
  alist(
    stem_length ~ dnorm(mu, sigma),
    mu <- a[genotype] + bT*treatment_s + bI[genotype]*treatment_s + bF[flat_factor],
    a[genotype] ~ dnorm(140, 30),
    bT ~ dnorm(0, 30),
    bI[genotype] ~ dnorm(0, 20),
    bF[flat_factor] ~ dnorm(0, 1),
    sigma ~ dexp(1)
  ), data=newphyE_list , chains=4 , cores=4 , log_lik=TRUE)

precis( modle_fixed, depth=2 )
```


#### c) use a hierarchical model that allows partial pooling across flats

```{r}
modle_multilevel <- ulam(
  alist(
    stem_length ~ dnorm(mu, sigma),
    mu <- a[genotype] + bT*treatment_s + bI[genotype]*treatment_s + bF[flat_factor],
    a[genotype] ~ dnorm(140, 30),
    bT ~ dnorm(0, 30),
    bI[genotype] ~ dnorm(0, 20),
    bF[flat_factor] ~ dnorm(f_bar, sigma_f),
    f_bar ~ dnorm(0, 1),
    sigma_f ~ dexp(1),
    sigma ~ dexp(1)
  ), data=newphyE_list , chains=4 , cores=4 , log_lik=TRUE)

precis( modle_multilevel, depth=2 )
```


### Q3) Compare the models, which is preferred?

```{r}
compare(modle_base, modle_fixed, modle_multilevel)
plot(compare(modle_base, modle_fixed, modle_multilevel))
```


### Q4) Using the hierarchical model, make posterior predictions
#### a) for average cluster

```{r}
## R code 13.32
post <- extract.samples(modle_multilevel)
str(post)

as.factor(newphyE_S_flat2$genotype)

#Shade condiction
## R code 13.36
p_link_abar_shade <- function( genotype , treatment_s=1) {
    stem_length_shade <- with( post, a[,genotype] + bT*treatment_s + bI[,genotype]*treatment_s )
    return( stem_length_shade )
}

## R code 13.37
p_raw_shade <- sapply( 1:6 , function(i) p_link_abar_shade( i ) )
p_mu_shade <- apply( p_raw_shade , 2 , mean )
p_ci_shade <- apply( p_raw_shade , 2 , PI )

plot( NULL , xlab="genotype" , ylab="stem_length_shade" ,
    ylim=c(0,260) , xaxt="n" , xlim=c(1,6) )
axis( 1 , at=1:6 , labels=c("Moneymaker", "phyB1", "phyB1/B2", "phyB2", "phyEami3", "phyEami7") )
lines( 1:6 , p_mu_shade )
shade( p_ci_shade , 1:6 )



#light condiction
## R code 13.36
p_link_abar_light <- function( genotype , treatment_s=0) {
    stem_length_light <- with( post, a[,genotype] + bT*treatment_s + bI[,genotype]*treatment_s )
    return( stem_length_light )
}

## R code 13.37
p_raw_light <- sapply( 1:6 , function(i) p_link_abar_light( i ) )
p_mu_light <- apply( p_raw_light , 2 , mean )
p_ci_light <- apply( p_raw_light , 2 , PI )

plot( NULL , xlab="genotype" , ylab="stem_length_light" ,
    ylim=c(0,260) , xaxt="n" , xlim=c(1,6) )
axis( 1 , at=1:6 , labels=c("Moneymaker", "phyB1", "phyB1/B2", "phyB2", "phyEami3", "phyEami7") )
lines( 1:6 , p_mu_light )
shade( p_ci_light , 1:6 )
```


#### b) for same clusters
```{r}

as.factor(newphyE_S_flat2$flat2)

#Shade condiction
## R code 13.36
p_link_abar_shade <- function( genotype , treatment_s=1, flat_factor=1) {
    stem_length_shade <- with( post, a[,genotype] + bT*treatment_s + bI[,genotype]*treatment_s + bF[,flat_factor] )
    return( stem_length_shade )
}

## R code 13.37
p_raw_shade <- sapply( 1:6 , function(i) p_link_abar_shade( i ) )
p_mu_shade <- apply( p_raw_shade , 2 , mean )
p_ci_shade <- apply( p_raw_shade , 2 , PI )

plot( NULL , main= "shade_1", xlab="genotype" , ylab="stem_length_shade" ,
    ylim=c(0,260) , xaxt="n" , xlim=c(1,6) )
axis( 1 , at=1:6 , labels=c("Moneymaker", "phyB1", "phyB1/B2", "phyB2", "phyEami3", "phyEami7") )
lines( 1:6 , p_mu_shade )
shade( p_ci_shade , 1:6 )



#light condiction
## R code 13.36
p_link_abar_light <- function( genotype , treatment_s=0, flat_factor=7) {
    stem_length_light <- with( post, a[,genotype] + bT*treatment_s + bI[,genotype]*treatment_s + bF[,flat_factor])
    return( stem_length_light )
}

## R code 13.37
p_raw_light <- sapply( 1:6 , function(i) p_link_abar_light( i ) )
p_mu_light <- apply( p_raw_light , 2 , mean )
p_ci_light <- apply( p_raw_light , 2 , PI )

plot( NULL , main= "sun_1", xlab="genotype" , ylab="stem_length_light" ,
    ylim=c(0,260) , xaxt="n" , xlim=c(1,6) )
axis( 1 , at=1:6 , labels=c("Moneymaker", "phyB1", "phyB1/B2", "phyB2", "phyEami3", "phyEami7") )
lines( 1:6 , p_mu_light )
shade( p_ci_light , 1:6 )

######2

#Shade condiction
## R code 13.36
p_link_abar_shade <- function( genotype , treatment_s=1, flat_factor=2) {
    stem_length_shade <- with( post, a[,genotype] + bT*treatment_s + bI[,genotype]*treatment_s + bF[,flat_factor] )
    return( stem_length_shade )
}

## R code 13.37
p_raw_shade <- sapply( 1:6 , function(i) p_link_abar_shade( i ) )
p_mu_shade <- apply( p_raw_shade , 2 , mean )
p_ci_shade <- apply( p_raw_shade , 2 , PI )

plot( NULL , main= "shade_2", xlab="genotype" , ylab="stem_length_shade" ,
    ylim=c(0,260) , xaxt="n" , xlim=c(1,6) )
axis( 1 , at=1:6 , labels=c("Moneymaker", "phyB1", "phyB1/B2", "phyB2", "phyEami3", "phyEami7") )
lines( 1:6 , p_mu_shade )
shade( p_ci_shade , 1:6 )



#light condiction
## R code 13.36
p_link_abar_light <- function( genotype , treatment_s=0, flat_factor=8) {
    stem_length_light <- with( post, a[,genotype] + bT*treatment_s + bI[,genotype]*treatment_s + bF[,flat_factor])
    return( stem_length_light )
}

## R code 13.37
p_raw_light <- sapply( 1:6 , function(i) p_link_abar_light( i ) )
p_mu_light <- apply( p_raw_light , 2 , mean )
p_ci_light <- apply( p_raw_light , 2 , PI )

plot( NULL , main= "sun_2", xlab="genotype" , ylab="stem_length_light" ,
    ylim=c(0,260) , xaxt="n" , xlim=c(1,6) )
axis( 1 , at=1:6 , labels=c("Moneymaker", "phyB1", "phyB1/B2", "phyB2", "phyEami3", "phyEami7") )
lines( 1:6 , p_mu_light )
shade( p_ci_light , 1:6 )

######3

#Shade condiction
## R code 13.36
p_link_abar_shade <- function( genotype , treatment_s=1, flat_factor=3) {
    stem_length_shade <- with( post, a[,genotype] + bT*treatment_s + bI[,genotype]*treatment_s + bF[,flat_factor] )
    return( stem_length_shade )
}

## R code 13.37
p_raw_shade <- sapply( 1:6 , function(i) p_link_abar_shade( i ) )
p_mu_shade <- apply( p_raw_shade , 2 , mean )
p_ci_shade <- apply( p_raw_shade , 2 , PI )

plot( NULL , main= "shade_3", xlab="genotype" , ylab="stem_length_shade" ,
    ylim=c(0,260) , xaxt="n" , xlim=c(1,6) )
axis( 1 , at=1:6 , labels=c("Moneymaker", "phyB1", "phyB1/B2", "phyB2", "phyEami3", "phyEami7") )
lines( 1:6 , p_mu_shade )
shade( p_ci_shade , 1:6 )



#light condiction
## R code 13.36
p_link_abar_light <- function( genotype , treatment_s=0, flat_factor=9) {
    stem_length_light <- with( post, a[,genotype] + bT*treatment_s + bI[,genotype]*treatment_s + bF[,flat_factor])
    return( stem_length_light )
}

## R code 13.37
p_raw_light <- sapply( 1:6 , function(i) p_link_abar_light( i ) )
p_mu_light <- apply( p_raw_light , 2 , mean )
p_ci_light <- apply( p_raw_light , 2 , PI )

plot( NULL , main= "sun_3", xlab="genotype" , ylab="stem_length_light" ,
    ylim=c(0,260) , xaxt="n" , xlim=c(1,6) )
axis( 1 , at=1:6 , labels=c("Moneymaker", "phyB1", "phyB1/B2", "phyB2", "phyEami3", "phyEami7") )
lines( 1:6 , p_mu_light )
shade( p_ci_light , 1:6 )
```

#### c) showing the "marginal" from cluster
```{r}
## R code 13.38
###shade
f_sim <- with( post , rnorm( length(post$f_bar) , f_bar , sigma_f ) )
p_link_fsim_shade <- function( genotype , treatment_s=1) {
    stem_length_shade <- with( post , f_sim + a[,genotype] + bT*treatment_s + bI[,genotype]*treatment_s )
    return( stem_length_shade )
}
p_raw_asim_shade <- sapply( 1:6 , function(i) p_link_fsim_shade( i ) )
p_mu_asim_shade <- apply( p_raw_asim_shade , 2 , mean )
p_ci_asim_shade <- apply( p_raw_asim_shade , 2 , PI )

plot( NULL , main= "shade_marginal", xlab="genotype" , ylab="stem_length_shade" ,
    ylim=c(0,260) , xaxt="n" , xlim=c(1,6) )
axis( 1 , at=1:6 , labels=c("Moneymaker", "phyB1", "phyB1/B2", "phyB2", "phyEami3", "phyEami7") )
lines( 1:6 , p_mu_asim_shade )
shade( p_ci_asim_shade , 1:6 )

###light
p_link_fsim_light <- function( genotype , treatment_s=0) {
    stem_length_light <- with( post , f_sim + a[,genotype] + bT*treatment_s + bI[,genotype]*treatment_s )
    return( stem_length_light )
}
p_raw_asim_light <- sapply( 1:6 , function(i) p_link_fsim_light( i ) )
p_mu_asim_light <- apply( p_raw_asim_light , 2 , mean )
p_ci_asim_light <- apply( p_raw_asim_light , 2 , PI )

plot( NULL , main= "sun_marginal", xlab="genotype" , ylab="stem_length_light" ,
    ylim=c(0,260) , xaxt="n" , xlim=c(1,6) )
axis( 1 , at=1:6 , labels=c("Moneymaker", "phyB1", "phyB1/B2", "phyB2", "phyEami3", "phyEami7") )
lines( 1:6 , p_mu_asim_light )
shade( p_ci_asim_light , 1:6 )
```


#### d) showing new clusters.
```{r}
## R code 13.39
###shade
plot( NULL , main= "new_clusters_shade", xlab="treatment" , ylab="stem_length_shade" ,
    ylim=c(0,260) , xaxt="n" , xlim=c(1,6) )
axis( 1 , at=1:6 , labels=c("Moneymaker", "phyB1", "phyB1/B2", "phyB2", "phyEami3", "phyEami7") )
for ( i in 1:100 ) lines( 1:6 , p_raw_asim_shade[i,] , col=col.alpha("black",0.25) , lwd=2 )

###sun
plot( NULL , main= "new_clusters_sun", xlab="treatment" , ylab="stem_length_light" ,
    ylim=c(0,260) , xaxt="n" , xlim=c(1,6) )
axis( 1 , at=1:6 , labels=c("Moneymaker", "phyB1", "phyB1/B2", "phyB2", "phyEami3", "phyEami7") )
for ( i in 1:100 ) lines( 1:6 , p_raw_asim_light[i,] , col=col.alpha("black",0.25) , lwd=2 )
```

### Q5) Reparameterize the model to help with divergent transitions (even if there aren't any)
```{r}
modle_multilevel <- ulam(
  alist(
    stem_length ~ dnorm(mu, sigma),
    mu <- a[genotype] + bT*treatment_s + bI[genotype]*treatment_s + bF[flat_factor],
    a[genotype] ~ dnorm(140, 30),
    bT ~ dnorm(0, 30),
    bI[genotype] ~ dnorm(0, 20),
    bF[flat_factor] ~ dnorm(f_bar, sigma_f),
    f_bar ~ dnorm(0, 1),
    sigma_f ~ dexp(1),
    sigma ~ dexp(1)
  ), data=newphyE_list , chains=4 , cores=4 , log_lik=TRUE)

precis( modle_multilevel, depth=2 )
```

### Q6--optional)
#### a) Which genotypes differ from MoneyMaker in Sun conditions?
```{r}

```

#### b) Which genotypes differ from MoneyMaker in Shade conditions?
```{r}

```

#### c) Which genotypes differ from MoneyMaker in their response to shade (difference in sun vs shade)?
```{r}

```

